%!TEX root = main.tex

\section{Instantiations of the generic decision procedure}\label{sec-instan}

In this section, we instantiate the generic decision procedure presented in the preceding section. In particular, the string functions of the assignments are modelled as parametric transducers. More precisely, each string function $f(x_1, \cdots, x_k)$ ($k\geq 1$) can be modelled by a parametric transducer $\Transducer$  
with $k-1$ parameters, and the size of $f$, $|f|$, is naturally defined by the size of $\Transducer$ (i.e., $|\Transducer|$). For the pre-image of an FA $\Aut$,  $\Pre_{R_f}(\Aut)$ is instantiated by $\Pre_\Transducer(\Aut)$, namely, the set of tuples $(w_1,\ldots, w_k)$ such that there is an accepting run of $\Transducer$ on $w_1$ with output $w'\in \Lang(\Aut)$ when equipped with the parameters $w_2,\ldots, w_k$. To apply Theorem~\ref{thm-generic-dec}, one only needs to verify the \prerec{} assumption. 
%While the latter trivially holds as we only consider assertions being regular constraints, the former is not trivial. 
For the rest of this section, we consider three classes of parametric transducers, viz. two-way transducers, one-way transducers, and reversal-bounded two-way transducers. We shall show that in each case, the pre-image is a recognisable relation with various size bounds $\ell(|\Transducer|, |\Aut|)$. This would yield tight complexity bounds on the path feasibility problem for string manipulating programs $S$ when the string functions therein are given in parametric transducers.  



%the set of tuples $(w, w_1,\ldots, w_k)$ such that there is an accepting run of $\Transducer$ on $w$ with output $w'\in \Lang(\Aut)$ when equipped with the parameters $w_1,\ldots, w_k$. 
 
\subsection{Two-way transducers}\label{sec-2way}

%Let $\Transducer$ be a 2NPT with $k$ parameters $y_1,\cdots, y_k$ and $\Aut$ be an NFA. Then the \emph{preimage} of $\Aut$ w.r.t. $\Transducer$, denoted by $\Pre_\Transducer(\Aut)$, is the set of tuples $(w, w_1,\cdots, w_k)$ such that there is an accepting run of $\Transducer$ on $w$ producing an output $w' \in \Lang(\Aut)$, equipped with the parameters $w_1,\cdots, w_k$.


\begin{lemma}\label{lem-2pt}
The \prerec{} assumption holds for string functions definable in \PPT{}s, with $\ell(|\Transducer|, |\Aut|) = 2^{\bigO(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$.
%Let $\Transducer$ be a \PPT{} and $\Aut$ be a conjunctive \FA{} $\Aut$. Then $\Pre_\Transducer(\Aut)$ is a recognisable relation. Moreover, a representation of $\Pre_\Transducer(\Aut)$, whose atom size is bounded by $2^{O(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$, can be computed effectively. Moreover, each disjunct of the representation can be nondeterministically computed in $2^{O(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$ space.
\end{lemma}

\begin{proof}
Assume a \PPT{} $\Transducer=(X, Q, q_0, F, \delta)$ with $X = \{x_1,\cdots, x_k\}$ and a conjunctive \FA{} $\Aut = ((Q', \delta'), S')$ with $S' \subseteq Q' \times Q'$. %We first introduce some notations. 
Let $\oalphabet = \ialphabet \cup X$.

To compute $\Pre_\Transducer(\Aut)$, we introduce a conjunctive \FFA{}\footnote{Conjunctive \FFA{}s are defined similarly to conjunctive \FA{}s.} $\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q}$, %defined as follows. 
%and $\Aut[S]$ defined below.
%
where  $q \in F$, and  $S_{x_1}, \cdots, S_{x_k} \subseteq Q' \times Q'$.   
The conjunctive \FFA{} $\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q}$ is constructed from $\Transducer$  by checking that there is an accepting run of $T$ on $w$, ending at the state $q$, and simulating an accepting run of $\Aut$ on the output of $\Transducer$. Therefore, each state of $\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q}$ is of the form $(q,q') \in Q \times Q'$. 
%In particular, 
When a transition of $\Transducer$ outputs a string $w'$ from $\oalphabet^*$ where some parameter $x_i$ appears (instead of a constant string), for each occurrence of $x_i$ in $w'$, a pair of states $(q_1', q_2')$ will be nondeterministically selected from $S_{x_i}$, and the second component of the state of $\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q}$ will be updated from $q_1'$ to $q_2'$ to mimic a possible instantiation of $x_i$.  

%\tl{zhilin, I am not quite clear about the para, $y_1$ is a typo of $x_1$? I wrote a para above; please check}
%(instead of materialising the output), where $S_{x_1}$ is used to jump over $y_1$ when $x_1$ is the output of a transition of $\Transducer$, similarly for $S_{x_2}, \cdots, S_{x_k}$. 


More precisely, $\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q}$ is computed as
%\begin{enumerate} 
%\item 
%the conjunctive \FFA{} 
$\Aut'' = ((\controls'', \transrel''), S'')$, where $Q'' = Q \times Q'$, $S'' = \{((q_0, p), (q, p')) \mid (p, p') \in S'\}$, and the transition relation $\transrel''$ comprises the tuples $((q_1, q'_1), a, dir, (q_2, q'_2))$ such that there is $w' \in \oalphabet^*$ satisfying that $(q_1, a, dir, q_2, w') \in \transrel$, and one of the following constraints holds, 
\begin{itemize}
\item $w' \in \ialphabet^*$  and $q'_1 \xrightarrow[\Aut]{w'} q'_2$, 
%
\item suppose $w' = w'_1 x_{i_1} \ldots w'_{r} x_{i_{r}} w'_{r+1}$ such that $r \ge 1$, $w'_j \in \ialphabet^*$ for each $j \in [r+1]$,  and $i_j \in [k]$ for each $j \in [r]$, then there are states $p_0, p_1, \ldots, p_{2r+1} \in \controls'$ satisfying that $p_0 = q'_1$, $p_{2r+1} = q'_2$, and 
$$p_0 \xrightarrow[\Aut]{w'_1} p_1 \xrightarrow{S_{x_{i_1}}} p_2\ \ldots\ p_{2r-2} \xrightarrow[\Aut]{w'_r}p_{2r-1} \xrightarrow{S_{x_{i_r}}} p_{2r} \xrightarrow[\Aut]{w'_{r+1}} p_{2r+1},$$ 
where $p_{2j-1} \xrightarrow{S_{x_{i_j}}} p_{2j}$ means $(p_{2j-1}, p_{2j}) \in S_{x_{i_j}}$ for each $j \in [r]$.
\end{itemize}
%
%
%\item Define $\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q}$ as $((Q''', \delta'''), S')$, where $S' = \{(\vec{\rho}_1, \vec{\rho}_2) \mid  (q'_1,q'_2) \in S, \vec{\rho}_1[1] =(q_0,q'_1), \vec{\rho}_2[|\vec{\rho}_2|] = (q, q'_2) \}$.
%\end{enumerate} 
It is easy to see that  the size of $\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q}$ is at most $|\Transducer||\Aut|$ and $|S''| = |S'|$.   
%$2^{O(|T||\Aut| \log(|T||\Aut|))}$.   

%For $S \subseteq Q' \times Q'$, let $\Aut[S]$ denote the product of $\Aut(q'_1, \{q'_2\})$ for $(q'_1,q'_2) \in S$ (cf. Section~\ref{sec:prelim}, {Operations of NFAs.}). Since $S$ contains at most $|Q'|^2=|\Aut|^2$ many elements, the size of $\Aut[S]$ is bounded by $|\Aut|^{|\Aut|^2} \approx 2^{O(|\Aut|^2 \log |\Aut|)}$.


\paragraph{Claim.} %For each tuple of strings $(w, w_1,\cdots, w_k)$, 
$(w, w_1,\cdots, w_k) \in \Pre_\Transducer(\Aut)$ iff there are $S_{x_1}, \cdots, S_{x_k} \subseteq Q' \times Q'$ and $q \in F$ such that $w \in \Lang(\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k}, q})$ and $w_i \in \Lang(\Aut[S_{x_i}])$ for each $i \in [k]$, where $\Aut[S_{x_i}]=((Q', \delta'), S_{x_i})$.

\medskip


Therefore, $\Pre_\Transducer(\Aut)$ is equal to 
\[
\bigcup_{S_{x_1}, \cdots, S_{x_k} \subseteq Q' \times Q', q\in F} \Lang(\cB_{\Transducer, \Aut, S_{x_1}, \cdots, S_{x_k},q}) \times \Lang(\Aut[S_{x_1}]) \times \cdots  \times \Lang(\Aut[S_{x_k}])\]

We conclude that $\Pre_\Transducer(\Aut)$ is a recognisable relation. 

In order to compute a representation of $\Pre_\Transducer(\Aut)$, we transform the conjunctive \FFA{} $\cB_{\Transducer, S_{x_1}, \cdots, S_{x_k},q}$ to a union of conjunctive \FA{}s as follows: Transform $\Aut''$ into a one-way transition graph $(\controls''',\delta''')$, where $\controls'''$ are vectors of states of $\Aut''$ of lengths at most $2|\controls''|-1$ (cf. Proposition~\ref{prop-2nfa-nfa}). Then $\Lang(\cB_{\Transducer, S_{x_1}, \cdots, S_{x_k},q})$ is the union of $\Lang(((\controls''',\delta'''), S''))$ for $S'' \subseteq \controls''' \times \controls''',$ which comprises nondeterministically selected pairs $(\vec{\rho}_1, \vec{\rho}_2) \in \controls''' \times \controls'''$, one for each $(p, p') \in S'$, such that $\vec{\rho}_1[1] =(q_0, p)$ and  $\vec{\rho}_2[|\vec{\rho}_2|] = (q, p')$.
%that is computed nondeterministically from $\cB_{\Transducer, S_{x_1}, \cdots, S_{x_k},q}$ and $S'$ as follows:  For each $(p, p') \in S$, nondeterministically select one pair $(\vec{\rho}_1, \vec{\rho}_2) \in \controls''' \times \controls'''$ satisfying $\vec{\rho}_1[1] =(q_0, p)$ and  $\vec{\rho}_2[|\vec{\rho}_2|] = (q, p')$, then put it into $S''$.

Since the size of each conjunctive \FA{} $((\controls''',\delta'''), S'')$ is $ 2^{\bigO(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$, we conclude that 
the \prerec{} assumption holds for string functions definable in \PPT{}s with $\ell(|\Transducer|, |\Aut|) = 2^{O(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$.
%$\Pre_\Transducer(\Aut)$ is a recognisable relation and a representation of $\Pre_\Transducer(\Aut)$, whose atom size is bounded by $2^{O(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$, can be constructed effectively.
\end{proof}

From Lemma~\ref{lem-2pt} and Theorem~\ref{thm-generic-dec}, by some simple calculation, we deduce that given a string manipulating program $S$ in SSA form where string functions are modelled by \PPT{}s, its path feasibility can be decided in 
%nondeterministic 
%$$O((\rcdim(\varphi)+2)^{\rcdep(\varphi)} (f^{\langle \rcdep(\varphi) \rangle}(|\varphi|, |\psi|))^2 \log f^{\langle \rcdep(\varphi) \rangle}(|\varphi|, |\psi|))$$ 
%space, where $f(j, k) = 2^{jk \log (j k)}$. 
nondeterministic $\tower(\rcdep(S), \bigO(\rcphi(S) \rcpsi(S) \log(\rcphi(S) \rcpsi(S)))$ space, thus in (deterministic) $\tower(\rcdep(S), \bigO(\rcphi(S) \rcpsi(S) \log(\rcphi(S) \rcpsi(S)))$ space, according to Savitch's theorem \cite{Savitch70}. Here the function $\tower(j, k)$ is %a tower of $2$s of height $j$ with $k$ on top, formally, 
defined as: $\tower(1, k)= 2^k$ and $\tower(j+1, k) = 2^{\tower(j, k)}$. We conclude that the path feasibility of such string manipulating programs  can be decided in $\rcdep(S)$-exponential space.
Moreover, we have a matching lower bound whose proof is sketched below.
%
\begin{theorem}
Given a string manipulating program $S$ in SSA form where string functions are modelled by \PPT{}s, the path feasibility of $S$ can be decided in $\rcdep(S)$-exponential space. Moreover,  for any $\expheight$, the path feasibility of such string manipulating programs is  %$\straightline[\twpt]$ 
 $\expheight$-$\expspace$-hard.
\end{theorem}

\paragraph{Lower-bound.}

We show that, for any $\expheight$, the problem is $\expheight$-$\expspace$-hard with $(\expheight+1)$ transducers, via a reduction from a tiling problem that is hard for $\expheight$-$\expspace$.
For this we need to manipulate large numbers that will be used to index the tiles in a solution to the tiling problem.
Similar encodings appear in the study of higher-order programs (e.g.~\cite{J01,CW07}) except the machinery needed to enforce the encoding is quite different.
The complete encoding is given in 
\shortlong{the full version of this article}
          {Appendix~\ref{sec:two-way-lower}}.
We give an overview of the intuition here.

A \emph{tiling problem} consists of a finite set of tiles \tiles as well as horizontal and vertical tiling relations 
$\hrel, \vrel \subseteq \tiles \times \tiles$.
Given a tiling \emph{corridor} of a certain width, as well as initial and final tiles
$\inittile, \fintile \in \tiles$
the task is to find a tiling where 
\begin{itemize}
\item
    the first tile of the first row of the corridor is $\inittile$, 
\item
    the tile $\tile$ directly to the left of $\tile'$ in a row must have
    $\tup{\tile, \tile'} \in \hrel$,
\item
    the tile $\tile$ in the row above $\tile'$ in the same column must have
    $\tup{\tile, \tile'} \in \vrel$,
\item
    the last position of the last row must contain $\fintile$.
\end{itemize}
Tiling problems are intimately connected to computations of Turing machines as
each row can hold the tape contents in one configuration of a run.
Thus, the corridor width is equivalent to the space of a Turing machine.

The existence of a solution to a tiling problem of width $n$ can be reduced to checking whether a \FFT\ of size 
$\ap{\bigoh}{n}$ 
can output a specified symbol $\top$.
Equivalently, we could use a \FFA.
A solution to a tiling problem will be encoded as a word 
$t_{1,1} t_{1,2} \cdots t_{1,n} \numsep \cdots \numsep t_{m,1} t_{m,2} \cdots t_{m,n}$
where each row
$t_{i,1} t_{i,2} \cdots t_{i,n}$
is separated from the next row by $\numsep$.
The \FFT\ performs $n+1$ passes.
During the first pass it checks that the tiling begins with $\inittile$, ends with $\fintile$ and we have 
$\tup{t_{i,j}, t_{i,j+1}} \in \hrel$
for all
$1 \leq j < n$.
If the encoding passes this check, we need to verify that the vertical tiling relation $\vrel$ is met.
This is done in $n$ more passes, the $j$th of which verifies the $j$th column of the encoding.

To encode tiling problems of $n$-fold exponential width we chain $(n+1)$ \FFT{}s.
We begin with two \FFT{}s and a tiling problem of exponential width.
Intuitively, a row of the solution to a tiling problem could now be encoded with a sequence of $n$ binary bits preceding each tile.
That is
\[
    0\ldots00\ \tile_{1,1}\ %
    0\ldots01\ \tile_{1,2}\ %
    \cdots
    1\ldots11\ \tile_{1,2^n}\ %
    \numsep
    \cdots
    \numsep
    0\ldots00\ \tile_{m,1}\ %
    0\ldots01\ \tile_{m,2}\ %
    \cdots
    1\ldots11\ \tile_{m,2^n} \ .
\]
In fact, in the formal encoding, we will not use binary numbers, but instead use tiling problems whose solutions must have exponential length.
For the purposes of the intuitive exposition, we will use binary numbers as the intention is clearer.

Given such an encoding, the first \FFT{} checks the solution in a similar manner to the width $n$ problem.
To cope with the large corridor width, it will output ``queries'' to the second \FFT.
These queries check whether one binary number is equal to the next, or is the increment.
Now, the first \FFT{} can, for example, non-deterministically guess that one tile is in the same column as another and use queries to verify this guess.
In this way it becomes able to verify a tiling over an exponentially wide corridor. \zhilin{Maybe say more explicitly as follows: when verifying the vertical tiling relation, for each column, the first \FFT{}  nondeterministically collects and outputs all the tiles and their addresses in this column, and ask the second \FFT{} to check the addresses of these tiles are equal and the vertical tiling relation holds in this column. The first \FFT{} goes through $2^n$ passes on the input, since there are $2^n$ columns. The second \FFT{} is also responsible for checking that after the tiles of the $i$-th column are outputted  by the first \FFT{}, then the tiles of the $(i+1)$-th column should be outputted. @Matt, is this description helpful here ?}

To increase the corridor width by another exponential, we need another \FFT.
To reach doubly-exponentially large numbers, we need to precede each tile with a binary sequence of exponential length.
For this we precede each bit with another binary sequence, this time of length $n$.
Thus, we have two sets of binary digits $0,1$ and $0', 1'$, and, if $n = 2$ a row would take the following form
\[
    (0'0') 0 (0'1') 0 (1'0') 0 (1'1') 0\ \tile_1\ %
    (0'0') 0 (0'1') 0 (1'0') 0 (1'1') 1\ \tile_2\ % 
    \ldots 
\]
where parenthesis are shown for clarity only.
To check a given tiling, the first \FFT\ outputs queries to the second.
To be able to check exponentially long binary numbers, the second \FFT\ outputs queries to the third \FFT.
This process can be iterated so that using $(n+1)$ \FFT, we can search for a solution of a tiling problem over an $n$-fold exponentially wide corridor.



%========================================one-way transducers==========================================

%\subsection{One-way transducers} \label{sec-one-way}


%\begin{lemma}\label{lem-1pt}
%The \prerec{} assumption holds for string functions definable in \PT{}s, with $\ell(|\Transducer|, |\Aut|) = |\Transducer||\Aut|$.
%%Let $\Transducer$ be a 1PT and $\Aut$ be a conjunctive NFA. Then $\Pre_\Transducer(\Aut)$ is a recognisable relation. Moreover, a representation of $\Pre_\Transducer(\Aut)$, whose atom size is bounded by $|\Transducer||\Aut|$, can be computed effectively.
%\end{lemma}

%Lemma~\ref{lem-1pt} can be proved in the same way as Lemma~\ref{lem-2pt}. The only difference is that the construction of $\cB_{\Transducer, \Aut, S_{x_1}, \ldots, S_{x_k}, q}$ is simpler and its size is at most $|\Transducer| |\Aut|$, since the transformation from  conjunctive \FFA{} to conjunctive \FA{} is not needed.

%From Lemma~\ref{lem-1pt} and Theorem~\ref{thm-generic-dec}, we deduce the following result.
%%
%\begin{theorem} \label{thm-1pt}
%%Given an $\straightline[\owpt]$ formula $\varphi \wedge \psi$, 
%The path feasibility problem of string manipulating programs in SSA form where string functions are modelled by \PT{}s is $\expspace$-complete. 
%\end{theorem}

%The lower bound in Theorem~\ref{thm-1pt} is deduced from the result in \cite{LB16} that the path feasibility of straight-line string constraints with concatenation and one-way transducers, which are both special cases of \PT{}s, is $\expspace$-complete.

%====================================

\subsection{$k$-Reversal-bounded two-way transducers}
 
%From Section \ref{sec-2way}  and Section \ref{sec-one-way}, we see that one-way transducers enjoy much lower complexity, but their expressiveness is limited. (For instance, the reverse function is inexpressible.) On the other hand, 

The result from Section \ref{sec-2way} reveals two-way transducers may suffer from very higher complexity. We observe that, in practice, many string functions, when encoded by transducers, need only a small number of reversals for the reading head, rather than the full power of two-way transducers. An extreme case is one-way transducers in which the reading head does not change the direction. 

%
%We shall consider reversal bounded transducers, which remain the complexity advantages of one-way transducers, but which can capture a larger class of string functions than one-way transducers. 

 
Let $k \in \Nat$. A \PPT{} $\Transducer$ is said  to be $k$-reversal bounded if in every run of $\Transducer$ on input strings, the reading head is reversed for at most $k$ times. We use  $k$-\RBPPT{} to denote the class of $k$-reversal bounded \PPT{}s. Note that the number of reversals in accepting runs of \PPT{}s must be even, since they must stop on the right end of input strings.
Evidently, the class of $0$-reversal bounded \PPT{}s is precisely the class of one-way \PT{}s.   
%
%As demonstrated in Example~\ref{exmp-ft} (in Section~\ref{sec:framework}), 
As another example, the string reverse function (cf. Example~\ref{exmp-ft} in Section~\ref{sec:framework}) can be captured by a $2$-reversal bounded \PPT.

%To handle $k$-reversal bounded transducers, 
An important  observation is  any string manipulating program (in SSA forms) with string functions modelled by general $k$-reversal bounded transducers can be translated to the one with only one-way transducers and a special transducer $\Transducer_{\sf reverse}$ for the reverse function.  

\begin{proposition} \label{prop:trans}
	Given a string manipulating program $S$ where  string functions are modelled by a $k$-reversal bounded transducer, one can construct $S'$ using only one-way transducers or  $\Transducer_{\sf reverse}$, and $ \rcdep(S')\in \mathcal{O}(k\cdot \rcdep(S))$. 
\end{proposition}

%Our goal is to show the following result.
%
To instantiate the generic procedure as in the preceding sections, one only needs to verify the  \prerec{} assumption for one-way transducers and $\Transducer_{\sf reverse}$. Indeed, although $\Transducer_{\sf reverse}$ is a two-way transducer, its pre-image does \emph{not} bring an exponential blowup as general two-way transducers would have (cf. Lemma~\ref{lem-2pt}), but rather it is akin to one-way transducers (cf. Lemma~\ref{lem-1pt}). 

\begin{lemma}\label{lem-1pt}
	The \prerec{} assumption holds for string functions definable in \PT{}s or $\Transducer_{\sf reverse}$, with $\ell(|\Transducer|, |\Aut|) = |\Transducer||\Aut|$.
	%Let $\Transducer$ be a 1PT and $\Aut$ be a conjunctive NFA. Then $\Pre_\Transducer(\Aut)$ is a recognisable relation. Moreover, a representation of $\Pre_\Transducer(\Aut)$, whose atom size is bounded by $|\Transducer||\Aut|$, can be computed effectively.
\end{lemma}

%\begin{lemma}\label{lem-reverse}
%	The \prerec{} assumption holds for  $\Transducer_{\sf reverse}$, with $\ell(| \Transducer_{\sf reverse}|, |\Aut|) = \bigO(|\Transducer_{\sf reverse}||\Aut|)$.
%	%Let $\Transducer$ be a \PPT{} and $\Aut$ be a conjunctive \FA{} $\Aut$. Then $\Pre_\Transducer(\Aut)$ is a recognisable relation. Moreover, a representation of $\Pre_\Transducer(\Aut)$, whose atom size is bounded by $2^{O(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$, can be computed effectively. Moreover, each disjunct of the representation can be nondeterministically computed in $2^{O(|\Transducer||\Aut|\log (|\Transducer||\Aut|))}$ space.
%\end{lemma}

Lemma~\ref{lem-1pt} can be proved in the same way as Lemma~\ref{lem-2pt}. The only difference is that the construction of $\cB_{\Transducer, \Aut, S_{x_1}, \ldots, S_{x_k}, q}$ is simpler and its size is at most $|\Transducer| |\Aut|$, since the transformation from  conjunctive \FFA{} to conjunctive \FA{} is not needed. For $\Transducer_{\sf reverse}$, to compute the pre-image one only needs to reverse the direction of each transition of $\Aut$ and swap the initial and final states.   

Combing Theorem~\ref{thm-generic-dec}, Proposition~\ref{prop:trans}, and Lemma~\ref{lem-1pt} together, we have:

\begin{theorem} \label{thm-rb-2pt}
%Given an $\straightline[\owpt]$ formula $\varphi \wedge \psi$, 
%For each constant $k \in \Nat$, the satisfiability problem of $\straightline[\rbtwpt_k]$ is $\expspace$-complete. 
For each $k \in \Nat$, the path feasibility of string manipulating programs in SSA form where string functions are modelled by $k$-\RBPPT{}s is $\expspace$-complete. 
\end{theorem}

The lower bound in Theorem~\ref{thm-rb-2pt} is deduced from the result in \cite{LB16} that the path feasibility of straight-line string constraints with concatenation and one-way transducers, which are both special cases of one-way \PT{}s, is already $\expspace$-hard.

%The $\expspace$-hardness in Theorem~\ref{thm-rb-2pt} follows from Theorem~\ref{thm-1pt}.
%The upper bound of Theorem~\ref{thm-rb-2pt} will not proved by directly instantiating Theorem~\ref{thm-generic-dec}, since  by doing this, we are only able to get a 2-$\expspace$ upper bound: Based on the construction in Section~\ref{sec-2way}, we can derive that the \prerec{} assumptions holds for string functions definable in $k$-\RBPPT, with $\ell(|\Transducer|, |\Aut|) = (|\Transducer||\Aut|)^{k+1}$. Nevertheless, $\ell^{\langle \rcdep(S)\rangle}(|\Transducer|, |\Aut|) = |\Transducer|^{(k+1) + \ldots + (k+1)^{\rcdep(S)}} |\Aut|^{(k+1)^{\rcdep(S)}}$, which is doubly exponential if $k > 0$.

%We prove the upper bound in Theorem~\ref{thm-rb-2pt} as follows: We first show that the $\expspace$ upper bound holds for the extension of \PT{}s with the reverse function. We then reduce in polynomial time the satisfiability problem for $k$-\RBPPT{}s to that for \PT{}s plus reverse function. The details can be found in Appendix~\ref{app-rb-2pt}. \zhilin{Matt, I wrote a very rough sketch of the reduction in appendix. Could you please help finish it ? This is a minor difference in the reduction. I do not use truncate transducer, instead, I add additional  indices to indicate the number of reversals that have been used. }

 
%We prove the upper bound in Theorem~\ref{thm-rb-2pt} as follows: We first show that the $\expspace$ upper bound holds for the extension of \PT{}s with the reverse function. We then reduce in polynomial time the satisfiability problem for $k$-\RBPPT{}s to that for \PT{}s plus reverse function. The details can be found in Appendix~\ref{app-rb-2pt}. 

\zhilin{Matt, could you write a proof in the appendix ?}
 
 


%===========================================================
\hide{
\begin{lemma}\label{lem-rb-2pt}
The \prerec{} assumptions holds for string functions definable in $k$-\RBPPT, with $\ell(|\Transducer|, |\Aut|) = (|\Transducer||\Aut|)^{k+1}$.
%Let $k \in \Nat$, $\Transducer$ be a RB2PT$_k$, and $\Aut$ be a conjunctive NFA. Then $\Pre_\Transducer(\Aut)$ is a recognisable relation. Moreover, a representation of $\Pre_\Transducer(\Aut)$, whose atom size is bounded by $(|\Transducer||\Aut|)^k$ \tl{k+1?}, can be computed effectively.
\end{lemma}

The proof of Lemma~\ref{lem-rb-2pt} is the same as Lemma~\ref{lem-2pt}. The only difference is that the size of $\cB_{\Transducer, \Aut, S_{x_1}, \ldots, S_{x_k}, q}$ is at most $(|\Transducer| |\Aut|)^k$, since $\Transducer$ is reversal $k$-bounded.

From Lemma~\ref{lem-rb-2pt} and Theorem~\ref{thm-generic-dec}, we deduce the following result.
%

The $\expspace$-hardness in Theorem~\ref{thm-rb-2pt} follows from Theorem~\ref{thm-1pt}.
}

%\input{expspace-hardness.tex}

%To show the above result, we will give a constraint that represents a straight-line program which ``transforms'' an automaton into exponentially many different automata.
%These exponentially many automata will be used to check the vertical tiling relation for a tiling problem with an exponentially wide corridor.
%Since such tiling problems are $\expspace$-hard, the result follows.

%\begin{theorem}\label{thm-1pt}
%	Satisfiability of $\strline[1PT]$, even in the single-letter case, is $\expspace$-hard.
%\end{theorem}


%
%\subsection{reversal-bounded transducers ?}
%
%\tl{include two-way non-parametric as well?}
