\documentclass{llncs}


%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{latexsym}
\usepackage{setspace}
\usepackage{cancel}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{appendix}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{leftidx}
\usepackage{mathtools}
\usepackage{paralist}
\usepackage{color}
\usepackage{mathrsfs}
\usepackage{tikz}
%\usepackage[draft]{minted}
\usetikzlibrary{shapes}
\usepackage[linesnumbered,ruled]{algorithm2e}

%==========================================================
\input{macros.tex}
\input{comm-anthony}
%==========================================================

%\newcommand\shortlong[2]{#2}
\newcommand\shortlong[2]{#1}

\newif\ifdraft\drafttrue
%\newif\ifdraft\draftfalse
\ifdraft
\newcommand{\anthony}[1]{\color{red} {YA: #1 :AY} \color{black}}
\newcommand{\zhilin}[1]{\color{brown} {ZL: #1 :LZ} \color{black}}
\newcommand{\tl}[1]{\color{blue} {TL: #1 :LT} \color{black}}
\newcommand{\mat}[1]{\color{cyan} {MH: #1 :HM} \color{black}}
\else
\newcommand{\anthony}[1]{}
\newcommand{\zhilin}[1]{}
\newcommand{\tl}[1]{}
\newcommand{\mat}[1]{}
\fi

\newcommand{\concat} {\circ}
\newcommand{\replace} {{\sf replace}}
\newcommand{\str} {{\sf Str}}
\newcommand{\intnum} {{\sf Int}}
\newcommand{\regexp} {{\sf RegExp}}
\newcommand{\strarr} {{\sf StringArray}}
\newcommand{\dtypes} {{\sf DataTypes}}
\newcommand{\anarr} {{\mathbb{A}}}

%============================================================


\begin{document}

\title{Parameterised Transducers}
\subtitle{(An Expressive Framework for Symbolic 
Execution Analysis of Programs with Strings)}

\author{}
\institute{}

\maketitle

\zhilin{Parametric versus parameterised ?}

\input{abstract}

\input{intro}

\input{prelim}

\input{framework}

\input{algorithms}

\input{extensions}

\input{conc}


\section{Parametric transducers}

In this section, we introduce parametric transducers.

\section{String constraints} \label{sec-core}

In this section, we define a general string constraint language that supports
concatenation, transducers, and regular constraints.
\tl{for length constraints, let's decide later whether it should be put here}
Throughout this section, we fix an alphabet $\Sigma$.
We consider the String data type $\str$, and assume a countable set of variables
$x, y, z, \cdots$ of $\str$.


\begin{definition}[Relational and regular constraints]
	Relational constraints and regular constraints are defined by the following rules,
	\[
	\begin{array}{r c l cr}
	s &\eqdef & x \mid u & \ \ & \mbox{(string terms)}\\
%	p &\eqdef & x \mid e & \ \ & \mbox{(pattern terms)}\\
	%t &\eqdef & s \mid e & \ \ & \mbox{(terms)}\\
	\varphi &\eqdef & x = s \concat s  \mid  x = T(\vec{s}) \mid \varphi \wedge \varphi & \ \ & \mbox{(relational constraints)}\\
	\psi & \eqdef & x \in e \mid \psi \wedge \psi %\mid \psi \vee \psi \mid \neg \psi
	& \ \ & \mbox{(regular constraints)} \\
	\end{array}
	\]
	where $x$ is a string variable, $u \in \Sigma^\ast$ and $e$ is a regular expression over $\Sigma$.
\end{definition}
\tl{this is not optimal. For $T$ with multiple parameters, the concatenation is redundant}

For a formula $\varphi$ (resp. $\psi$), let $\vars(\varphi)$ (resp. $\vars(\psi)$) denote the set of variables occurring in $\varphi$ (resp. $\psi$). Given a relational constraint $\varphi$, a variable $x$ is called a \emph{source variable} of $\varphi$ if $\varphi$ \emph{does not} contain a conjunct of the form $x = s_1 \concat s_2$ or $x = T(\vec{s})$.

%We then notice that, with the $\replaceall$ function in its general form, the concatenation operation is in fact redundant.

%\begin{proposition}\label{prop-concat}
%	The concatenation  operation ($\concat$) can be simulated  by the $\replaceall$ function.
%\end{proposition}
%\begin{proof}
%	It is sufficient to observe that %the concatenation operator $s_1 \concat s_2$ is redundant in the sense that
%	a relational constraint $x = s_1 \concat s_2$ can be rewritten as
%	\[x' = \replaceall(ab, a, s_1) \wedge x = \replaceall(x', b, s_2),\] where $a,b$ are two fresh letters.
%\end{proof}

%In light of Proposition~\ref{prop-concat}, in the sequel, we will \emph{dispense the concatenation operator} mostly and focus on \textbf{the string constraints that involve  the $\replaceall$ function only}.

%Another example to show the power of the $\replaceall$ function is that it can simulate the extension of regular expressions with string variables, which is  supported by the mainstream scripting languages like Python, Javascript, and PHP. For instance, $x \in y^*$ can be expressed by $x =\replaceall(x', a, y) \wedge x' \in a^*$, where $x'$ is a fresh variable and $a$ is a fresh letter.



The generality of the constraint language makes it undecidable,
even in very simple cases. To retain decidability, we follow \cite{LB16} and focus on the ``straight-line fragment" of the language. This straight-line fragment captures the structure of straight-line string-manipulating
programs.

\begin{definition}[Straight-line relational constraints]
	A relational constraint $ \varphi$ with transducers is straight-line, if $\varphi \eqdef \bigwedge \limits_{1 \le i \le m} x_i = P_i$ such that
	\begin{itemize}
		\item $x_1,\dots, x_m$ are mutually distinct,
		\item for each $i \in [m]$, all the variables in $P_i$ are either source variables, or variables from $\{x_1,\dots, x_{i-1}\}$,
	\end{itemize}
	%Occasionally we refer to $x_m$ as the output variable.
\end{definition}
%Intuitively, in a straight-line relational constraint, the dependency graph (see Definition~\ref{def:dep-graph}) of the string variables is acyclic.
%\mat{forward reference!}

\begin{remark}
	Checking whether a relational constraint $\varphi$ is straight-line can be done in linear time.
\end{remark}

\begin{definition}[Straight-line string constraints]
	A straight-line string constraint $C$ with transducers (denoted by $\strline[T]$)  is defined as $ \varphi \wedge \psi$,  where
	\begin{itemize}
		\item $\varphi$ is a straight-line relational constraint with transducers,  and
		%
		\item $\psi$ is a regular constraint.
		%
	\end{itemize}
\end{definition}



 We first introduce a graphical representation of $\strline[T]$ formulae as follows.

 \begin{definition}[Dependency graph]
	\label{def:dep-graph}
	Suppose $C= \varphi \wedge \psi$ is an $\strline[\replaceall]$ formula where the pattern parameters of the $\replaceall$ terms are regular expressions. %Let $\vars(\varphi) = \{x_1,\dots, x_m, y_1, \dots, y_n\}$, where $y_1,\dots, y_n$ are  source variables.
	Define the \emph{dependency graph} of $C$ as $G_C= (\vars(\varphi), E_C)$, such that for each $i \in [m]$, if $x_i = \replaceall(z, e_i, z')$, then $(x_i, (\rpleft, e_i), z) \in E_C$ and $(x_i, (\rpright, e_i), z') \in E_C$. A final (resp. initial) vertex in $G_C$ is a vertex in $G_C$ without successors (resp. predecessors). The edges labelled by $(\rpleft, e_i)$ and $(\rpright, e_i)$ are called the $\rpleft$-edges and $\rpright$-edges respectively. The \emph{depth} of $G_C$ is the maximum length of the paths in $G_C$. In particular, if $\varphi$ is empty, then the depth of $G_C$ is zero.
	%The $\rpleft$-length of a path $\pi$, denoted by $\rpleftlen(\pi)$, is the number of $\rpleft$-edges on $\pi$. A path of $G_C$ is a sequence $z_1 \ell_1 z_2 \dots \ell_{k-1} z_k$ such that for each $i \in [k-1]$, $(z_i, \ell_i, z_{i+1}) \in E_C$. A path is initial (resp. final) if the path starts from an initial vertex (resp. stops at a final vertex).
	% e the $\src$-nesting-depth of $z$ in $G_C$, denoted by $\srcnd_{G_C}(z)$,  as the maximum number of $\src$-edges in paths from source variables to $z$.
 \end{definition}
 Note that $G_C$ is a DAG where the out-degree of each vertex is two or zero.

\subsection{The satisfiability problem} \label{sec-sat}
In this paper, we focus on the satisfiability problem of $\strline[T]$, which is formalised as follows.

%\smallskip

\begin{quote} \centering
	\framebox{Given an $\strline[T]$ constraint $C$, decide whether $C$ is satisfiable.}
\end{quote}
\smallskip

%To approach this problem, we identify several fragments of  $\strlineTT]$, depending on whether the pattern and the replacement parameters are constants or variables.  We shall investigate extensively the satisfiability problem of the fragments of $\strline[\replaceall]$. % (see Table~\ref{tab-sum}).  Note that for $x=\replaceall (y, p, z)$, $p$ is referred to as a \emph{pattern} and $z$ is referred to as a \emph{replacement}.

%=========================================================================================================

\begin{proposition}
undecidable in general
\end{proposition}

\section{A generic decision procedure}

\zhilin{results for two-way versus one-way, with parameters or without.}

In this section, we focus on the case where the transducer $T$ has only one parameter, i.e., the relational constraint is of form $x=T(y)$ for example. We will give both upper and low bounds.

\tl{A question: here there are two ways to present the result, one is to follow \cite{LB16} to encode $\concat$, and the other is to deal with $\concat$ explicitly. Which way do you think is better?}

\subsection{Upper-bound}
We first show that, when the 2-way transducer $T$ and an NFA


\subsection{Lower-bound}

\input{two-way-lower.tex}

%==========================================================================================

\section{Two-way transducers with length constraints}

This section is dynamical: we hope for the best of Anothy's result; in case it does not work, we have two possible backups: (1) reversal-bounded 2-way transducers; (2) using reversal-bounded counter machines to represent (both regular and length) constraints

%===========================================================================================

\section{One-way transducers with variables}

Not quite sure whether we need this section, it might be just a simple generalisation of the popl'18 paper, or be subsumed by the next section ; we will see.

%========================================================================================

\section{Two-way transducers with variables}

\subsection{Pre-image computation of 2-way transducers}

Let $\vec{y}=\{y_1, \cdots, y_m\}$.

The general idea is to encode a general string manipulating function $f(x, \vec{y})$ as a NFT $T$ over $\Sigma$ and $\Sigma\cup\{\vec{y}\}$. The question for the pre-image computation is formalised as follows:
\begin{itemize}
	\item INPUT: A NFT $T$, a regular language $\mathcal{A}$.
	\item OUTPUT: $(L^{(0)}_i, L^{(1)}_i, \cdots, L^{(m)}_i )_{i=1}^\ell$, such that
	\[\exists z\in\mathcal{A} \wedge z=f(x, \vec{y})\mbox{ iff }\exists k. x\in L^{(0)}_k \wedge y_i\in L^{(i)}_k \]
\end{itemize}

%===========================================================================================

%\section{Matt's pet :-), maybe another paper}
%
%\begin{definition}[multi-tape transducer, with k input tapes, and one output tape.]
%	\tl{Matt, please elaborate}
%	The
%	k input tapes follow a stack discipline: tape i can only move if the
%	head position of all tapes $j > i$ is 0
%\end{definition}


\section{Conclusion}

%==============================================================================================
\newpage

% Bibliography
\bibliographystyle{plain}
\bibliography{string}

\appendix

\end{document}
