%!TEX root = main.tex

\section{Symbolic extensions}
\label{sec:symbolic}

In this section, we consider an extension of string manipulating programs to the symbolic setting, where finite-state automata and parametric transducers are replaced by \emph{symbolic automata} \cite{NG01,DV14} and \emph{symbolic parametric transducers} (to be introduced) respectively.
%
Intuitively, in symbolic automata %are an extension of finite-state automata in which 
input letters are symbolically described by unary predicates from a logical theory. %On the other hand, 
%
Likewise, in symbolic parametric transducers %are an extension of parametric transducers in which 
output letters are given by terms in the logic theory and the parameters as in parametric transducers may occur.
%,  apart from the symbolic input letters. 
Symbolic parametric transducers can be seen as extensions of both the parametric
transducer introduced in this paper and the symbolic transducer introduced in
\cite{symbolic-transducer}.
%In the sequel, we first define symbolic automata and transducers, then introduce symbolic parametric transducers.
% and show how to extend the results in preceding sections to the symbolic setting.


%Let $\data$ be an infinite data domain.  We define data strings as elements of $\data^\ast$.
We assume familiarity with the %(many-sorted) 
first-order logic (cf.\ some textbook e.g. \cite{EFT94}). A concept of background theories is used has been in the definition of symbolic automata and transducers\footnote{In literature, a symbolic automata is defined on a Boolean algebra, which is a background theory without requiring the set of formulae to be closed under substitutions. We define symbolic automata on background theories (rather than Boolean algebras) in Definition~\ref{def-2sa} to align to the symbolic parametric transducers defined later.}.
%and label theories. 
Intuitively, 
%a Boolean algebra is a set of unary first-order logic formulae closed under Boolean connectives, 
a background theory is a set of unary first-order logic formulae closed under Boolean connectives and substitutions. 
 %Label theories extend background theories further by adding inequalities of terms into the set of formulae.
 
 
 \begin{definition}[Background theories]
A background theory $\Upsilon$ is a tuple $(\signature, \structure, \Psi)$ satisfying the following constraints:
\begin{itemize}
\item $\signature=(\functions, \predicates)$ is a signature, where $\functions$ (resp. $\predicates$) is a recursively enumerable set of function symbols (resp. relation symbols), 
%\begin{itemize}
%\item
%
%\item for each $\vec{s} = s_1 \times \dots \times s_n$ and $i \in [n]$, there is a function $\pi_{\vec{s}, i}$ of arity $s_1 \times \dots \times s_n \rightarrow s_i$ in $\functions$ such that $\pi_{\vec{s},i}(\vec{t}) = t_i$ for each term $\vec{t} = (t_1,\dots, t_n)$ of sort $s_1 \times \dots \times s_n$.
%\end{itemize}

\item $\structure$ is an $\signature$-structure $(\data, \interpretation)$, where $\data$ is a countably infinite set (the data domain), and $\interpretation$ is a function on $\data$ satisfying: for each $n$-ary function symbol $f \in \functions$ (resp. relation symbol $R \in \predicates$), $\interpretation(f)$ is an $n$-ary function on $\data$ (resp. $\interpretation(R)$ is an $n$-ary relation on $\data$),
%
\item $\Psi$ is a recursively enumerable set of \emph{unary} $\signature$-formulae closed under Boolean connectives $\vee, \wedge, \neg$ and substitutions, that is, for each $\signature$-term $t(x)$ and $\psi(x) \in \Psi$, we have $\psi[t(x)/x] \in \Psi$. 
%In addition, it is assumed that each formula $\psi \in \Psi$ of arity $s^\ell$ (where $\ell > 0$) has the same set of free variables $\{x_1,\dots, x_\ell\}$.
\end{itemize}
%A background theory is a Boolean algebra $(\signature, \structure, \Psi)$ such that $\Psi$ is closed under substitutions, that is, for each $\signature$-term $t(x)$ and $\psi(x) \in \Psi$, we have $\psi[t(x)/x] \in \Psi$. \\
%
For each $\psi(x) \in \Psi$, we use $\|\psi\|$ to denote the set $\{d \in \data \mid  \structure \models \psi(d)\}$. Elements of $\| \psi\|$ are called the \emph{witnesses} of $\psi$.
A formula $\psi \in \Psi$ is \emph{satisfiable}, denoted by $\issat(\psi)$, if $\|\psi\| \neq \emptyset$. A background theory $\Upsilon = (\signature, \structure, \Psi)$ is \emph{decidable} if  checking  $\issat(\psi)$ for a  given $\psi \in\Psi$ is decidable.
\end{definition}


In the rest of this section, we {\bf fix a decidable background theory} $\Upsilon=(\signature, \structure, \Psi)$ with $\signature=(\functions, \predicates)$ and $\structure = (\data, \interpretation)$. Moreover, we assume that given a formula $\psi \in \Psi$ of size $m$ in $\Upsilon$, $\issat(\psi)$ can be decided in $\bigO(\beta(m))$ space for some monotone function $\beta$ such that $\beta(m) \ge \log m$.
A \emph{data string} is an element of $\data^*$. A \emph{data language} is a set of data strings.



%%%%===========================================================
\hide{
\begin{definition}[Background theories]
A background theory  $\Upsilon$ is a tuple $(\signature, \structure, \Psi)$ satisfying the following constraints:
\begin{itemize}
\item $\signature=(\functions, \predicates)$ is a signature, where $\functions$ (resp. $\predicates$) is a recursively enumerable set of function symbols (resp. relation symbols), 
%\begin{itemize}
%\item
%
%\item for each $\vec{s} = s_1 \times \dots \times s_n$ and $i \in [n]$, there is a function $\pi_{\vec{s}, i}$ of arity $s_1 \times \dots \times s_n \rightarrow s_i$ in $\functions$ such that $\pi_{\vec{s},i}(\vec{t}) = t_i$ for each term $\vec{t} = (t_1,\dots, t_n)$ of sort $s_1 \times \dots \times s_n$.
%\end{itemize}

\item $\structure$ is an $\signature$-structure $(\data, \interpretation)$, where $\data$ is a countably infinite set (the data domain), and $\interpretation$ is a function on $\data$ satisfying: for each $n$-ary function symbol $f \in \functions$ (resp. relation symbol $R \in \predicates$), $\interpretation(f)$ is an $n$-ary function on $\data$ (resp. $\interpretation(R)$ is an $n$-ary relation on $\data$),
%
\item $\Psi$ is a recursively enumerable set of unary $\signature$-formulae closed under Boolean connectives $\vee, \wedge, \neg$. In addition, $\Psi$ is closed under substitutions, that is, for each $\signature$-term $t(x)$ and $\psi(x) \in \Psi$, we have $\psi[t(x)/x] \in \Psi$.
%In addition, it is assumed that each formula $\psi \in \Psi$ of arity $s^\ell$ (where $\ell > 0$) has the same set of free variables $\{x_1,\dots, x_\ell\}$.
For each $\psi(x) \in \Psi$, we use $\|\psi\|$ to denote the set $\{d \in \data \mid  \structure \models \psi(d)\}$. Elements of $\| \psi\|$ are called the \emph{witnesses} of $\psi$.
\end{itemize}
A formula $\psi \in \Psi$ is \emph{satisfiable}, denoted by $\issat(\psi)$, if $\|\psi\| \neq \emptyset$. In addition, $\Upsilon$ is \emph{decidable} if  checking  $\issat(\psi)$ for a  given $\psi \in\Psi$ is decidable.
\end{definition}
}
%%%%===========================================================


\begin{definition}[Symbolic automata]\label{def-2sa}
    A (nondeterministic two-way)  symbolic automata (\SSA) is a tuple $\Aut = (\EndLeft, \EndRight, \controls, q_0, \finals, \transrel)$, where  
\begin{itemize}
%\item $\Upsilon=(\signature, \structure, \Psi)$ is a decidable Boolean algebra, where $\signature = (\functions, \predicates)$ and $\structure = (\data, \interpretation)$,
%
\item $\EndLeft$, $\EndRight$, $\controls$, $q_0$, $\finals$ are defined precisely as in \FFA{}s, 
%
\item $\transrel$ is a finite set of transitions of one of the following forms,
\begin{itemize}
\item     symbolic transitions $(q, \psi(x), dir, q') \in \controls \times \Psi \times \{\Left, \Stay, \Right\} \times \controls$, where $q, q'$ are the source and target state, $dir$ is the direction, and $\psi(x)$ is the guard of the transition, 
%
\item     non-symbolic transitions $(q, \EndLeft, dir, q')$ with $dir \in \{\Right, \Stay\}$, or $(q, \EndRight, dir', q')$ with $dir' \in \{\Left,\Stay\}$. 
\end{itemize}
\end{itemize}
A \SSA{} is an \SA{} if there are no transitions with the direction ``$\Left$". 
%for each $(q, \psi(x), dir, q')$, we have $dir \in \{\Stay, \Right\}$, moreover, the transitions of the form $(q, \EndRight, \Left, q')$ are removed.
\end{definition}

\paragraph{Semantics of \SSA{} $\Aut = (\Upsilon, \EndLeft, \EndRight, \controls, q_0, \finals, \transrel)$.}
A transition $\tau=(q_1,\psi(x), dir, q_2) \in \transrel$ in the \SSA{}  can be concretised
into a potentially infinite set of \emph{concrete} transitions $\|\tau\| \subseteq Q \times \data \times \{\Left, \Stay, \Right\} \times Q$, where $(q_1, d, dir, q_2)  \in \|\tau\|$ iff $d \in \|\psi\|$.
Intuitively, suppose $\Aut$ is at the state $q_1$ and reading the input data value $d \in \data$.
If there is a transition $(q_1, \psi(x), dir, q_2) \in \transrel$ such that $d \in \|\psi\|$, then $\Aut$ can move its reading head according to $dir$, and change the state from $q_1$ to $q_2$.

%$\Aut$ is interpreted on \emph{data strings}, namely, elements of $\data^*$.
Given a data string $w = d_1, \dots, d_n$, a \emph{run} of $\Aut$ on $w$
is a sequence of tuples $(q_0, i_0), \ldots, (q_m, i_m) \in \controls \times [0, n+1]$ 
such that
%let $d_0 = \EndLeft$ and $d_{n+1} = \EndRight$, %. The following conditions, then, have to be satisfied:
\begin{itemize}
    \item $i_0 = 0$, and
    \item for every $j \in [m-1]$, one of the following holds,
    \begin{itemize}
  	\item $0 < i_j < n+1$,  and $\tau=(q_j, \psi(x), dir, q_{j+1}) \in \transrel$ for some $\psi(x) \in \Psi$ and $dir \in \{\Left, \Stay, \Right\}$, such that $(q_j, d_{i_j}, dir, q_{j+1}) \in \|\tau\|$, and $i_{j+1} = i_j + dir$,
	%
	\item $i_j = 0$, $(q_j, \EndLeft, dir, q_{j+1}) \in \transrel$, and $i_{j+1} = i_j + dir$,
	%
	\item $i_j = n+1$, $(q_j, \EndRight, dir, q_{j+1}) \in \transrel$, and $i_{j+1} = i_j + dir$.
  \end{itemize}
\end{itemize}
The run is said to be \defn{accepting} if $i_m = n+1$ and $q_m \in \finals$.
A data string $w$ is accepted by $\Aut$ if there is an accepting run of
$\Aut$ on $w$. We use $\Lang(\Aut)$ to denote the data language recognised by $\Aut$, that is,  the set of data strings accepted by $\Aut$.
%
For simplicity, we usually omit $\EndLeft$, and $\EndRight$ from the definition of \SSA{}s. 

To measure the size of a \SSA{} $\Aut$, we need to take the guards of the transitions (from the background theory $\Upsilon$) into  consideration. Therefore, we assume that the formulae in $\Psi$ are encoded in a proper way. %e.g. encoded in binary. 
For $\psi \in \Psi$, we use $|\psi|$ to denote the number of symbols in the encoding of $\psi$. Moreover, we define the size of a symbolic transition $\tau = (q_j, \psi(x), dir, q_{j+1})$ to be $|\psi|$.
For a refined complexity analysis, we use the following two measures of $\Aut$: %numbers to capture the size of $\Transducer$:  Define 
the \emph{structural size} and the \emph{data size} of $\Aut$, denoted by $|\Aut|_s$ and $|\Aut|_d$ respectively, where $|\Aut|_s$ is defined as the number of transitions of $\Aut$, and $|\Aut|_d$ is defined as the maximum size of symbolic transitions of $\Aut$.


\begin{proposition}\label{prop-2nsa}
Assume a \SSA{}  $\Aut$. Then
\begin{itemize}
\item $\Aut$ can be  transformed into an equivalent \SA{} $\Aut'$ in time $|\Aut'|_s \cdot |\Aut'|_d$, where $|\Aut'|_s = 2^{\bigO\left(|\Aut|_s \log |\Aut|_s\right)}$ and $|\Aut'|_d = \bigO(|\Aut|_s |\Aut|_d)$;
%$|\Aut'| = 2^{O (|\Aut| \log |\Aut| )}$.
%
%
%\item SAs are under the intersection operator\footnote{As a matter of fact, SAs are closed under all Boolean operations. Since we only use the intersection in this paper, we restrict our attention to the intersection.}. \zhilin{complexity}
%
\item the nonemptiness of $\Aut$ can be decided in nondeterministic $\bigO( (\log |\Aut|_c) + \beta(|\Aut|_d))$ space.
\end{itemize}
\end{proposition}
The proof of the first result in Proposition~\ref{prop-2nsa} can be found in Appendix~\ref{app-sym}, while the second result is obtained by guessing a path in $\Aut$ and was shown in \cite{NG01,DV14}.

%\smallskip

%Next, we introduce symbolic transducers.


%we derive a new sort $s^*$ such that $\|s^* \| = \bigcup \limits_{i \in \Nat}\data_s^i$, where $\data^i_s = \{\varepsilon\}$. An element of $\|s^* \|$ is called a \emph{data string} of sort $s$.




% ======= label theories are removed, which are only useful for deterministic SA ================
% ======= label theories are removed, which are only useful for deterministic SA ================
\hide{
\begin{definition}[Label theories]
A label theory  $\Upsilon$ is a tuple $(\signature, \structure, \Psi, \Psi')$ satisfying the following constraints:
\begin{itemize}
\item $(\signature, \structure, \Psi)$ is a background theory with $\signature = (\functions, \predicates)$,
%
\item $\Psi'$ is the set of formulae of the form $\psi(x) \wedge t_1(x) \neq t_2(x)$, where $\psi(x) \in \Psi$ and $t_1(x), t_2(x)$ are $\signature$-terms.
\end{itemize}
A label theory is decidable if it is decidable to check $\issat(\psi)$ for $\psi \in \Psi \cup \Psi'$.
\end{definition}

Given a formula $\psi(x) \in \Psi$ and two $\signature$-terms $t_1(x), t_2(x)$,
$t_1$ and $t_2$ are \emph{equivalent up to $\psi$}, denoted by $t_1 \simeq_\psi t_2$, if
 $\issat(\psi(x) \wedge t_1(x) \neq t_2(x))$ does not hold.
%Two sequences of $s^i/ s$-terms $\vec{f}=f_1...f_n$ and $\vec{g}=g_1...g_m$ are \emph{equivalent up to $\psi$}, denoted by $\vec{f}\simeq_\psi \vec{g}$,
%iff $n=m$ and for every $j \in [n]$, $f_j \simeq_\psi g_j$.

%Given an $s^i/ s$-term sequence $\vec{f}=f_1...f_n$ and a sequence of data values $\vec{d} = (d_1, \dots, d_i) \in (\data_{s_{\sf in}})^i$,
%let $\|\vec{f}\|(\vec{d})$ denote the sequence $\|f_1\|(\vec{d})...\|f_n\|(\vec{d})$, that is, a data string of sort $s$.
}
% ======= label theories are removed, which are only useful for deterministic SA ================
% ======= label theories are removed, which are only useful for deterministic SA ================


\begin{definition}[Symbolic finite-state transducers]
    A (nondeterministic two-way) symbolic transducer (\SST) is an extension of an \SSA{} with outputs. More precisely, a \SST{} $\Transducer$ is a tuple $(\EndLeft, \EndRight, \controls, q_0, \finals, \transrel)$ where  
\begin{itemize}
%\item $\Upsilon=(\signature, \structure, \Psi)$ is a decidable background theory, where $\signature = (\functions, \predicates)$ and $\structure = (\data, \interpretation)$,
%
\item $\EndLeft$, $\EndRight$, $\controls$, $q_0$, $\finals$ are defined precisely as in \SSA{}s, 
%
\item $\transrel$ is a finite set of  transitions which are in one of the following forms,
\begin{itemize}
\item  symbolic transitions $(q, \psi(x), dir, q', \vec{t}(x))$ such that $q, q' \in Q$, $dir \in \{\Left, \Stay, \Right\}$, $\psi(x) \in \Psi$, and
$\vec{t}(x) = t_1(x) \ldots t_r(x)$ is a (possibly empty) sequence of $\signature$-terms, 
\item   $(q, \EndLeft, dir, q', \epsilon)$ or $(q, \EndRight, dir', q', \epsilon)$, with $dir \in \{\Right, \Stay\}$ and $dir' \in \{\Left, \Stay\}$,
\end{itemize}
\end{itemize}
A \SST{} is an \ST{} if there are no transitions with the direction ``$\Left$". 
%$(q, \psi(x), dir, q', \vec{t}(x))$ we have $dir \in \{\Stay, \Right\}$ and the transitions of the form $(q, \EndRight, \Left, q', c)$ are not present.
\end{definition}

\paragraph{Semantics of \SST{}.}
The semantics of \SST{}s is defined similarly to that of \SSA{}s. 
A transition $\tau=(q_1,\psi(x), dir, q_2, \vec{t}(x)) \in \transrel$ with $\vec{t}(x) = t_1(x) \ldots t_r(x)$ in the \SST{} $\Transducer$ can be concretised
into a potentially infinite set of \emph{concrete} transitions $\|\tau\| \subseteq Q \times \data \times \{\Left, \Stay, \Right\} \times Q \times \data^r$, where $(q_1, d, dir, q_2, d'_1\ldots d'_r)  \in \|\tau\|$ iff $d \in \|\psi\|$ and $d'_j = \interpretation(t_j)(d)$ for each $j \in [r]$.
%Intuitively, suppose $\Transducer$ is at the state $q_1$ and reading the input data value $d \in \data$.
%If there is a transition $(q_1, \psi(x), dir, q_2, t(x) )\in \transrel$ and $d \in \|\psi\|$, $\Transducer$ can move its reading head according to $dir$, update the current state $q_1$ to $q_2$,  and produce the data value $d' \in \data$.

Given a data string $w = d_1, \dots, d_n$, a \emph{run} of $\Transducer$ on $w$
is a sequence of tuples $(q_0, i_0, \vec{d}'_0), \ldots, (q_m, i_m, \vec{d}'_m) \in \controls \times [0, n+1] \times \data^*$ 
such that
%let $d_0 = \EndLeft$ and $d_{n+1} = \EndRight$, %. The following conditions, then, have to be satisfied:
\begin{itemize}
    \item $i_0 = 0$, and
    \item for every $j \in [m-1]$, one of the following holds,
    \begin{itemize}
  	\item  $0< i_j < n+1$, $\tau=(q_j, \psi(x), dir, q_{j+1}, \vec{t}(x)) \in \transrel$ for some $\psi(x) \in \Psi$, $dir \in \{\Left, \Stay, \Right\}$, and a sequence of  $\signature$-terms $\vec{t}(x)$, such that $(q_j, d_{i_j}, dir, q_{j+1}, \vec{d}'_j) \in \|\tau\|$, and $i_{j+1} = i_j + dir$,
	%
	\item $i_j = 0$, $(q_j, \EndLeft, \Right, q_{j+1}, \epsilon) \in \transrel$, and $\vec{d}'_j  = \epsilon$,
	%
	\item $i_j = n+1$, $(q_j, \EndRight, \Left, q_{j+1}, \epsilon) \in \transrel$, and $\vec{d}'_j  = \epsilon$.
  \end{itemize}
\end{itemize}
The run is said to be \defn{accepting} if $i_m = n+1$ and $q_m \in \finals$. When a run is accepting, $\vec{d}'_0, \ldots, \vec{d}'_m$ is the \emph{output} of the run.
A data string $w'$ is said to be an output of $\Transducer$ on $w$ if there is an accepting run of
$\Transducer$ on $w$ with output $w'$. We use $\Tran(\Transducer)$ to denote the \emph{transduction} defined by $\Transducer$, that is, the relation comprising the data-string pairs $(w, w')$ such that $w'$ is an output of $\Transducer$ on $w$.
%
We define $|\Transducer|_s$ and $|\Transducer|_d$, the structural and data size of $\Transducer$, by taking the outputs of transitions into consideration. Specifically, $|\Transducer|_s$ is defined as the product of the number of transitions and the maximum length of sequences of $\signature$-terms in  transitions of $\Transducer$; 
$|\Transducer|_d$ is defined as the maximum size of symbolic transitions of $\Transducer$, where the size of a symbolic transition $(q, \psi(x), dir, q', \vec{t}(x))$ with $\vec{t}=t_1 \ldots t_r$ is defined as $|\psi| + |t_1| + \ldots + |t_r|$.
%the number of transitions of $\Transducer$, and $\sizetrans(\Transducer)$, the maximum size of transitions of $\Transducer$.

We are ready to introduce symbolic parametric transducers.

\begin{definition}[Symbolic parametric transducers]
A nondeterministic two-way symbolic parametric transducer (\SSPT) is an extension of a \SST{} with parameters. More precisely,  a \SSPT{} $\Transducer$ is a tuple
$(\EndLeft, \EndRight, Y, \controls, q_0, \finals, \transrel)$ where:
\begin{itemize}
%\item $\Upsilon=(\signature,\structure, \Psi, \Psi')$ is a decidable background theory with $\signature = (\functions, \predicates)$ and $\structure = (\data, \interpretation)$,
%
\item $\EndLeft$, $\EndRight$, $\controls$, $q_0$, $\finals$ are defined precisely as in \SST{}s, 
%
\item $Y=\{y_1,\ldots, y_m\}$ is a finite set of parameters, 
%
\item $\transrel$ is a finite set of transitions in one of the following forms: 
\begin{itemize}
\item symbolic transitions $(q, \psi(x), q', \vec{t})$ such that $q,q' \in Q$, $\psi(x) \in \Psi$,
$\vec{t} = t_1 \ldots t_r$ such that for each $i \in [r]$, either $t_i$ is an $\signature$-term where only the variable $x$ occurs, or $t_i \in Y$,
%
\item non-symbolic transitions $(q, \EndLeft, dir, q', \epsilon)$, $(q, \EndRight, dir, q', \epsilon)$, with $dir \in \{\Right, \Stay\}$ and $dir' \in \{\Left, \Stay\}$. 
\end{itemize}
\end{itemize}
\end{definition}

Notice that, as in \PPT{}s, parameters are only allowed in the output track.
Intuitively, each instantiation of the parameters $y_1,\ldots, y_m$ with data strings 
$w_1,\ldots, w_m$ gives rise to a \SST{} $\Transducer[w_1/y_1,\ldots, w_m/y_m]$ by replacing each occurrence of $y_i$ with $w_i$. 
%outputs the data string $w_j$, whenever a transition of the form $(p, \psi, dir, q, y_j)$ is
%taken by the transducer. 
%This instantiation of the parameters is only done 
%\emph{once} before the symbolic parametric transducer is run.
%
We use $\transet(\Transducer)$ to denote the set of tuples $(w, w_1, \ldots, w_m, w')$ such that $(w, w') \in \transet(\Transducer[w_1/y_1,\ldots, w_m/y_m])$.
The two size measures of \SSPT{}s can be defined in a similar way as in \SST{}s. 


%%%============================================
%%%============================================
\hide{
%To extend the string constraint language $\straightline[\transet]$ to the symbolic setting, 
To lift the result to the symbolic setting, we only need  \emph{symbolic automata}, the counterpart of \FA. We start with \emph{Boolean algebra}, 
\tl{I am not an expert, but technically say $\Psi$ is a boolean algebra? More precisely, \SA{} works not on a background theory, but some more relaxed thing.}
which is adapted from a background theory $\Upsilon=(\signature,\structure, \Psi)$ by dropping out the requirement that $\Psi$ is closed under substitutions.

\tl{as we donot have $\straightlinesym[\transet]$ anymore? to be updated}
For a class $\transet$ of symbolic parametric transducers, we use $\straightlinesym[\transet]$ to denote the extension of straight-line string constraint language by replacing parametric transducers with symbolic parametric transducers and finite-state automata with symbolic automata.
}
%%%============================================
%%%============================================


\paragraph{Symbolic extension of the generic decision procedure.} We consider the path feasibility problem of \emph{data} string manipulating programs in SSA form 
%\tl{do we need to introduce formally?}
 where each assertion is a conjunction of atomic constraints of the form $x \in \Aut$ with $\Aut$ being an \SA. 
To solve the path feasibility problem for data string manipulating programs,
we generalise the generic decision procedure in Section~\ref{sec:algo} by replacing recognisable relations and the \prerec{} assumption with \emph{symbolic recognisable relations} and the \emph{$\mathbb{S}$\prerec{} assumption}.

\begin{definition}[Symbolic recognisable relations]
%	Given a decidable Boolean algebra $\Upsilon=(\signature, \structure, \Psi)$ with $\signature = (\functions, \predicates)$ and $\structure = (\data, \interpretation)$, 
	A $k$-ary relation $R \subseteq \data^*\times \ldots\times \data^*$ is a \emph{symbolic recognisable} relation if $R=\bigcup_{i=1}^n L^{(i)}_1 \times \ldots \times L^{(i)}_k$ where $L^{(i)}_j$ is recognised by some \SA{} for each $j\in [k]$.
\end{definition}
Similar to the recognisable relation, we also represent a  $k$-ary symbolic recognisable relation as a collection of tuples $(\Aut_1, \ldots, \Aut_k)$, where each atom $\Aut_i$ is a conjunctive representation of \SA{}, namely, of the form $((\controls, \transrel), S)$ with $S \subseteq \controls \times \controls$. We will use a conjunctive \SA{} to mean a conjunctive representation of the \SA{}. For a conjunctive \SA{} $\cC= ((\controls, \transrel), S)$, two size measures $|\cC|_s$ and $|\cC|_d$ are defined similarly to \SA{}s. Moreover, the structural (resp. data) size of a representation of a symbolic recognisable relation is defined as the maximum structural (resp. data) size of the atoms.

\begin{quote}
{\bf The $\mathbb{S}$\prerec{} assumption}. For each data string function $f$ in $S$ and each conjunctive \SA{} $\Aut$,  $\Pre_{R_f}(\Aut)$ is a symbolic recognisable relation. Furthermore, 
a representation of $\Pre_{R_f}(\Aut)$, whose structural size (resp. data size) is bounded by  $\ell_s(|f|_s, |\Aut|_c)$ (resp. $\ell_d(|f|_s,  |f|_d, |\Aut|_s, |\Aut|_d)$) for some monotone functions $\ell_s$ and $\ell_d$, can be computed effectively. We also assume that each disjunct of the representation can be nondeterministically computed in $\ell_s(|f|_s, |\Aut|_s) \cdot \ell_d(|f|_s,  |f|_d, |\Aut|_s,  |\Aut|_d)$ space.
%the space to compute a representation of $\Pre_{R_f}(\Aut)$ and the atom size of such a representation\footnote{Usually the amount of space is at least as large as the size of the representation; here for simplicity we assume that they are asymptotically the same.} are both bounded by $\ell(|f|, |\Aut|)$ for some monotone function $\ell$.  
%
%\item[A2] Each assertion $g$ in $S$ is given as a conjunction of atomic regular constraint $x\in \Aut$. 
\end{quote} 
%
Here $|f|_s$ and $|f|_d$ are the structural and data sizes of a representation of $f$ respectively; the concrete definitions depend on the form of $f$, which will be given when the generic decision procedure is instantiated later. Furthermore, we define the size of $f$ (resp. $|\Aut|$), denoted by $|f|$ (resp. $\Aut$), as $|f|_s \cdot |f|_d$ (resp. $|\Aut|_s \cdot |\Aut|_d$).

Let $S$ be a data string manipulating program in SSA form.  We use $|S|$ to denote the sum of $|f|$ and $|\Aut|$ for data string functions $f$ and \SA{}s $\Aut$ occurring in $S$.  
For a refined complexity analysis, we also define $\rcdep(S)$ and $\rcdim(S)$ similarly to the string setting. Moreover, we use $\rcphi_s(S)$ (resp. $\rcphi_d(S)$) to denote the maximum of $|f|_s$ (resp. $|f|_d$) for data string functions $f$ occurring in the assignments of $S$, and $\rcpsi_s(S)$ (resp. $\rcpsi_d(S)$) to denote the maximum of $|\Aut|_s$ (resp. $|\Aut|_d$) for \SA{}s $\Aut$ occurring in $S$. Finally, we define $\rcsreg(S)$ to be the maximum number of atomic \SA{} constraints for one variable in $S$.


%Similarly, we can define $\numtrans(\psi)$ and $\sizetrans(\psi)$.


%We can adapt the definition of $\numtrans$ and $\sizetrans$ to the conjunctive representations of SA easily. We will use conjunctive SA to mean a conjunctive representation of the SA.

For a binary function $g(k, l)$, $h(i, j, k, l)$, and $n \in \Nat \backslash \{0\}$, we define $g^{\langle n \rangle}(k, l)$ and $h^{\langle n \rangle}_g(i, j, k, l)$ as follows: $g^{\langle 1 \rangle}( k, l) = g(k, l)$, $h^{\langle 1 \rangle}_g(i, j, k, l) = h(i, j, k, l)$, and 
$$g^{\langle n+1 \rangle}( k, l) = g(k, g^{\langle n \rangle}( k, l)), \ h^{\langle n+1 \rangle}_g (i, j, k, l) = h(i, j, g^{\langle n \rangle}(k, l), h^{\langle n \rangle}(i, j, k, l)).$$


\begin{theorem}\label{thm-generic-dec-symbolic}
	Given a data string manipulating program $S$ satisfying the $\mathbb{S}$\prerec{} assumption, the path feasibility problem of $S$ can be decided in \emph{nondeterministic} $M+ \beta(M)$ space, where  
	\[
	\begin{array}{l c l}
		M & = & |\vars(S)| \cdot (\rcdim(S)+1)^{\rcdep(S)}  \rcsreg(S) \cdot  (\ell_s^{\langle \rcdep(S) \rangle}(\rcphi_s(S), \rcpsi_s(S)))^{r}\ \cdot \\
		& &  \hspace{1cm} (\ell_d)^{\langle  \rcdep(S) \rangle}_{\ell_c}(\rcphi_s(S),\rcphi_d(S),  \rcpsi_s(S), \rcpsi_d(S))
\end{array}
	\]
for some constant $r > 0$. 
\end{theorem}



%%%=============================================================
\hide{
\begin{theorem}\label{thm-generic-dec-symbolic}
Let $\transet$ be a class of symbolic parametric transducers. Suppose that %$\transet$ satisfies that 
for each $\Transducer \in \transet$ and conjunctive \SA{} $\Aut$, $\Pre_\Transducer(\Aut)$ is a symbolic recognisable relation and a representation of which can be computed effectively, 
such that for every atom $\Aut'$ therein, 
$\numtrans(\Aut') = O(g(\numtrans(\Transducer),  \numtrans(\Aut)))$ and 
$\sizetrans(\Aut') = O(h(\numtrans(\Transducer), \sizetrans(\Transducer), \numtrans(\Aut), \sizetrans(\Aut)))$, for two monotone functions $g$ and $h$.
Then the path feasibility of a $\straightlinesym[\transet]$ formula $\varphi \wedge \psi$ can be decided in nondeterministic space
$$
O\left(
\begin{array}{l}
(\rcdim(\varphi)+2)^{\rcdep(\varphi)}|\psi| \cdot (\numtrans'')^2 \log \numtrans'' + \\
f((\rcdim(\varphi)+2)^{\rcdep(\varphi)} (\numtrans'')^2 \sizetrans'')  
\end{array}
\right),
$$ 
where $\numtrans'' = g^{\langle \rcdep(\varphi) \rangle}(\numtrans(\varphi), \numtrans(\psi))$ and 
$$\sizetrans'' = h^{\langle \rcdep(\varphi) \rangle}_g(\numtrans(\Transducer), \sizetrans(\Transducer), \numtrans(\Aut), \sizetrans(\Aut)).$$
\end{theorem}
}
%%%=============================================================

\paragraph{Instantiation of the generic decision procedure to \SSPT{}s and \SPT{}s.}



%\noindent Finally, we instantiate the generic decision procedure to \SSPT{}s.

\begin{lemma}\label{lem-spt}
The $\mathbb{S}$\prerec{} assumption holds for \SSPT{}s with $\ell_s(|\Transducer|_s, |\Aut|_s) = 2^{\bigO( |\Transducer|_s |\Aut|_s^{|\Transducer|_s} \log (|\Transducer|_s |\Aut|_s^{|\Transducer|_s}))}$ and  $\ell_d(|\Transducer|_s, |\Transducer|_d, |\Aut|_s, |\Aut|_d) = \bigO(|\Transducer|_s  |\Transducer|^2_d |\Aut|_s^{|\Transducer|_s} |\Aut|_d)$. 

The $\mathbb{S}$\prerec{} assumption holds for \SPT{}s, with $\ell_s(|\Transducer|_s, |\Aut|_s) =\bigO(|\Transducer|_s |\Aut|_s^{|\Transducer|_s})$ and  $\ell_d(|\Transducer|_s, |\Transducer|_d, |\Aut|_s, |\Aut|_d) = \bigO(|\Transducer|^2_d |\Aut|_d)$.
%
%Given a \SSPT{} $\Transducer$ and a conjunctive \SA{} $\Aut$, $\Pre_\Transducer(\Aut)$ is a symbolic recognisable relation and a representation of which can be computed effectively, such that for each atom $\Aut'$ therein, 
%$|\Aut'|_c = 2^{\bigO(|\Transducer|_c |\Aut|_c \log (|\Transducer|_c |\Aut|_c))},$ 
%and 
%$|\Aut'|_d = |\Transducer|_c |\Aut|_c |\Transducer|_d |\Aut|_d.$
%On the other hand, if $\Transducer$ is an \SPT, then $|\Aut'|_c$ and $|\Aut'|_d$ are reduced to $\bigO(|\Transducer|_c |\Aut|_c)$ and $\bigO(|\Transducer|_d |\Aut|_d)$ respectively.
\end{lemma}


From Theorem~\ref{thm-generic-dec-symbolic} and Lemma~\ref{lem-spt}, we have the following result.
\begin{theorem}\label{thm-spt}
The path feasibility of data string manipulating programs can be decided 

\begin{itemize}
	\item in nondeterministic $\tower(\rcdep(S)+1, \bigO(|S|^2)) + \beta(\tower(\rcdep(S)+1, \bigO(|S|^2)))$  space if the data string functions are given by \SSPT{}s;
	
	\item  in nondeterministic $|S|^{|S|^{\bigO(k\cdot \rcdep(S))}} + \beta(|S|^{|S|^{\bigO(k\cdot \rcdep(S))}})$ space if the data string functions are given by $k$-RB2SPTs. %\SPT{}s;
	
\end{itemize}
%
%where 
%
%The path feasibility for \SPT{}s can be decided
\end{theorem}
%Note that for illustration purpose, we choose not to instantiate the decision procedure to  here.
%Similar instantiations can be done for $k$-RB2SPTs along the same line as in Section~\ref{subsec:krb}, which we omit for simplicity. 


%==========================================================================
%==========================================================================
%============================many sorted first-order logic=========================
%==========================================================================
%==========================================================================

\hide
{

\paragraph{Many-sorted first-order logic.}
We assume a signature $\signature=(\sorts, \functions, \predicates)$, where $\sorts$ is a countable set of \emph{sorts}, $\functions$ is a countable set of \emph{function symbols}, and $\predicates$ is a countable set of \emph{predicate symbols}. Each function or predicate symbol has an associated \emph{arity}, which is a tuple of sorts in $\sorts$.  A function symbol with a single sort is called a \emph{constant}. A predicate symbol with a single sort is called a \emph{set}, which intuitively denotes a set of elements of that sort.

An $\signature$-term is built from the function symbols in $\functions$ and variables taken from a set $\mathcal{X}$ that is disjoint from $\sorts$, $\functions$, and $\predicates$. Each variable $x \in \mathcal{X}$ has an associated sort in $\sorts$. In addition, we assume that the variables in $\mathcal{X}$ are linearly ordered $\preceq_{\mathcal{X}}$. When writing $t(\vec{x})$ for a vector of distinct variables $\vec{x}$ such that $\vec{x} = (x_1,\dots, x_n)$ follows the ascending order of the linear order $\preceq_{\mathcal{X}}$, we assume that the variables occurring in the term $t$ are from $\vec{x}$. For a term $t(\vec{x})$ of sort $s$ such that $\vec{x} = (x_1, \dots, x_n)$ and each $x_i$ for $i \in [n]$ is of sort $s_i \in \sorts$, the term $t$ is said to be \emph{of arity} $(s_1 \times \dots \times s_n) \rightarrow s$. In addition, for a vector of terms $(t_1, \dots, t_m)$ such that all the variables of $t_1 ,\dots, t_m$ are from $\vec{x} = (x_1, \dots, x_n)$, if $x_1 \preceq_{\mathcal{X}} x_2  \preceq_{\mathcal{X}} \dots  \preceq_{\mathcal{X}} x_n$, each $x_i$ for $i \in [n]$ is of sort $s_i$, and each $t_j$ for $j \in [m]$ is of sort $s'_j$, then $(t_1,\dots, t_m)$ is said to be a term of arity $(s_1,\dots, s_n) \rightarrow (s'_1,\dots, s'_m)$. For readability, a term of arity $(s_1,\dots, s_n) \rightarrow (s'_1,\dots, s'_m)$ is also called a $(s_1,\dots, s_n) \big/ (s'_1,\dots, s'_m)$-term. We use $(t_1,\dots, t_m)(\vec{x})$ to denote a vector of terms whose variables are from $\vec{x}$.  For convenience, we also write $t(\vec{x})$ as $\lambda \vec{x}.\ t$ and $(t_1,\dots, t_m)(\vec{x})$ as $\lambda \vec{x}.\ (t_1,\dots, t_m)$. 

We assume the standard notions of $\signature$-atoms, $\signature$-literals, and $\signature$-formulae, whose definitions can be found in some textbooks on mathematical logic (see e.g. \cite{Gal85}). The set of free variables of a $\signature$-formula $\psi$ is denoted by $\free(\psi)$. When writing $\psi(\vec{x})$, we assume that the free variables of $\psi$ are from $\vec{x}$. For a formula  $\psi(\vec{x})$ such that $\vec{x} = (x_1, \dots, x_n)$ and each $x_i$ for $i \in [n]$ is of sort $s_i \in \sorts$, the formula $\psi$ is said to be \emph{of arity} $s_1 \times \dots \times s_n$.
A formula $\psi$ that contains exactly one free variable (resp. two, $n \ge 3$ free variables) is called a \emph{unary} (resp. \emph{binary}, $n$-ary) $\signature$-formula. A formula $\psi$ contains no free variables is called a $0$-ary formula, aka a sentence. For $i, j \in \Nat \backslash \{0\}$, a formula $\psi(\vec{x})$ of arity $s^j$ (where $\vec{x}=(x_1, \dots, x_j)$), and an $s^i/s^j$-term $\vec{f}=(f_1,\dots, f_j)$, we use $\psi[\vec{f}/\vec{x}]$ to denote the formula obtained from $\psi$ by simultaneously replacing $x_1$ with $f_1$, $\dots$, and $x_j$ with $f_j$.

An $\signature$-interpretation $I$ maps: (i) each sort $s \in \sorts$  to a set $s^{I}$, (ii) each function symbol $f \in \functions$ of arity $s_1 \times \ldots \times s_n \rightarrow s$ to a total function $f^I: s_1^I \times \ldots \times s^I_n \rightarrow s^I$ if $n>0$, and to an element of $s^I$ if $n = 0$, and (iii) each predicate symbol $p \in \predicates$ of sort $s_1 \times \ldots \times s_n$ to a  subset of $p^I \subseteq s^I_1 \times \ldots s^I_n$.
An $\signature$-assignment $\eta$ maps each variable $x \in \mathcal{X}$ of sort $s \in \sorts$ to an element of $s^I$.
\begin{itemize}
\item For a term $t$, the interpretation of $t$ under $(I, \eta)$ for an $\signature$-interpretation $I$ and $\signature$-assignment $\eta$, denoted by $t^{(I,\eta)}$, can be defined inductively on the syntax of terms.
\item The satisfiability relation between pairs of an $\signature$-interpretation and an $\signature$-assignment, and $\signature$-formulae, written $I \models_{\eta} \psi$,
is defined inductively, as usual.
\end{itemize}
We say that $(I,\eta)$ is a model of $\psi$ if $I \models_{\eta} \psi$. For an $\signature$-sentence $\psi$, we also write $I \models \psi$ if there is an $\signature$-assignment $\eta$ such that $I \models_\eta \psi$.

Let $\signature$ be a signature and $\cI$ be a set of $\signature$-interpretations. Then $\theory(\cI)$, \emph{the $\signature$-theory associated with $\cI$}, is the set of  $\signature$-sentences $\psi$ such that for each $I \in \cI$, $I \models \psi$.


}
