\documentclass{llncs}
\usepackage{minted}
%\documentclass[envcountsame, fleqn]{llncs}
  \usepackage{amsmath, amssymb, latexsym}
  \usepackage{graphicx}
\usepackage{epic,eepic}
\usepackage{etoolbox}
%\pagestyle{plain}
\usepackage{listings}
\usepackage{psfrag}
\usepackage{rotating}

\usepackage{url}
\usepackage{amssymb,amsthm, epsfig,amstext}
\usepackage{txfonts}
\usepackage{algorithmic}
\usepackage[linesnumbered,ruled,noend]{algorithm2e}
\usepackage{graphicx}

\usepackage{mathrsfs}

\usepackage{color, colortbl}

\renewcommand{\floatpagefraction}{.8}%


\setminted{numbersep=5pt, xleftmargin=12pt}
\usemintedstyle{friendly} 

\usepackage{todonotes}
\usepackage{multirow}
\usepackage{float,color}
\usepackage{picinpar,color,xcolor,wrapfig}
\setcounter{secnumdepth}{4}
\sloppy

\pagestyle{plain}

\newtoggle{full}
%\toggletrue{full}
\togglefalse{full}

%%%%%%%%%%%%%%%   MACROS  %%%%%%%%%%%%%%%%%%


\input{macros.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\zhilin}[1]{\color{cyan} {ZL: #1 :LZ} \color{black}}
\newcommand{\tl}[1]{\color{purple} {TL: #1 :LT} \color{black}}
\newcommand{\anthony}[1]{\color{blue} {AL: #1 :LA} \color{black}}
\newcommand{\matt}[1]{\color{blue} {MH: #1 :HM} \color{black}}
\newcommand{\philipp}[1]{\color{blue} {PR: #1 :RP} \color{black}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\title{A Decision Procedure for Path Feasibility of String \\
 Manipulating Programs  with Integer Data Type}


\author{Taolue Chen\inst{1} \and Matthew Hague\inst{2} \and Jinlong He\inst{3,6} \and Denghang Hu\inst{3,6} \and \\ 
	Anthony Widjaja Lin\inst{4} \and Philipp R\"ummer\inst{5} \and Zhilin Wu\inst{3, 7, 8}}

\institute{University of Surrey, UK
	\and Royal Holloway, University of London, UK
	\and State Key Laboratory of Computer Science, \\
	Institute of Software, Chinese Academy of Sciences, China
	\and  Technical University of Kaiserslautern, Germany
    \and Uppsala University, Sweden
    \and University of Chinese Academy of Sciences, China
    \and Shanghai Key Laboratory of Trustworthy Computing, East China Normal University, China
    \and Institute of Intelligent Software, Guangzhou, China
}

\begin{document}
\maketitle

\vspace{-5mm}

\begin{abstract}
\input{abstract.tex}
\end{abstract}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-3mm}
\section{Introduction} \label{sec:intro}

%\vspace{-2mm}

\input{intro.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-1mm}
\section{Preliminaries}\label{sec:prel}

\input{prel.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-1mm}
\section{String-Manipulating Programs with Integer Data Type}\label{sec:logic}

\input{logic.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-1mm}
\section{Decision Procedures for Path Feasibility}\label{sec:dec}

\input{decision.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-5mm}
\section{Evaluations} \label{sec:eva}

\input{eval.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-3mm}
\section{Conclusion} \label{sec:conc}

In this paper, we have proposed  an expressive string constraint language which can specify constraints on both strings and integers.  We provided an automata-theoretic decision procedure for the path feasibility problem of this language. The decision procedure is simple, generic, and amenable to implementation, giving rise to a new solver OSTRICH+.  We have evaluated OSTRICH+ on  a wide range of existing and newly created benchmarks, and have obtained very encouraging results.  OSTRICH+ is shown to be the first solver  which is capable of tackling finite transducers and integer constraints with completeness guarantees. Meanwhile, it demonstrates competitive performance against some of the best state-of-the-art string constraint solvers.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\medskip
\small{
\noindent \emph{Acknowledgements.}  
T.~Chen and Z.~Wu are supported by Guangdong Science and Technology Department grant (No.\ 2018B010107004); T. Chen is also supported by Overseas Grant (KFKT2018A16) from the State Key Laboratory of
Novel Software Technology, Nanjing University, China and Natural Science Foundation of Guangdong Province, China (No. 2019A1515011689).
M.~Hague is supported by EPSRC [EP/T00021X/1];.
A.~Lin is supported by the European Research Council (ERC) under the European
Union's Horizon 2020 research and innovation programme (grant agreement no
759969).
P.~R\"ummer is supported by the
Swedish Research Council (VR) under grant 2018-04727, and by the
Swedish Foundation for Strategic Research (SSF) under the project
WebSec (Ref.\ RIT17-0011).
Z.~Wu is supported by  the Open Project of Shanghai Key Laboratory of Trustworthy Computing (No. 07dz22304201601), the NSFC grants (No. 61872340), and the INRIA-CAS joint research project VIP.   
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-3mm}
\bibliographystyle{abbrv}
\bibliography{string-short}

%\iffalse
%\newpage
%\setcounter{page}{1}

\iftoggle{full}
{
\newpage
\begin{appendix}
\input{appendix}
\end{appendix}
}
{}%\fi

\end{document}
