
\section{Introduction}



Strings are a fundamental data type in virtually all programming languages.
Their generic nature can, however, lead to many subtle programming
bugs, some with security consequences, e.g., cross-site scripting
(XSS), which is among the OWASP Top 10 Application Security Risks
\cite{owasp17}. 

One effective
automatic testing method for identifying subtle programming errors
%automatically generating test cases
%generation with a good coverage
is based on \emph{symbolic execution}
\cite{king76} and combinations with dynamic analysis
called \emph{dynamic symbolic execution} \cite{jalangi,DART,EXE,CUTE,KLEE}.
See \cite{symbex-survey} for an excellent survey. Unlike purely random testing,
which runs only \emph{concrete} program executions on different
inputs, the techniques of symbolic execution analyse \emph{static} paths
(also called symbolic executions) through the software system under test.
%Since such a path is simply a sequence of assignments and
%conditionals/assertions,
Such a path can be viewed as a constraint $\varphi$ (over
appropriate data domains) and the hope is that a fast
solver is available for checking the satisfiability of $\varphi$ (i.e. to check
the \emph{feasibility} of the static path), which can be used for generating
inputs that lead to certain parts of the program or an erroneous behaviour.
%a undesirable program behaviour.
%or an exploration of a new part of the
%system.


%
In this paper, we focus on two string operations with emphasis on extended regular expressions. 

s the \emph{string-replace function}, 
which may be used to replace either the \emph{first} occurrence or
\emph{all} occurrences of a string (a string constant/variable, or a regular expression) by 
another string (a string constant/variable). The replace function (especially 
the replace-all functionality) is omnipresent in HTML5 applications
\cite{LB16,TCJ16,YABI14}. 
%\mat{What does it mean for the replace function to be convincingly argued?}
For example, a standard industry defense against cross-site scripting 
(XSS) vulnerabilities includes sanitising untrusted strings before adding them
into the DOM (Document Object Model) or the HTML document. 
This is typically done by %replacing every occurrence of
various metacharacter-escaping mechanisms (see, for instance, 
\cite{Kern14,BEK,OWASP-XSS}). An example of such a mechanism is backslash-escape, which replaces \emph{every
	occurrence} of quotes and double-quotes (i.e. \verb+'+ and \verb+"+) in the
string by \verb+\'+ and \verb+\"+. 
In addition to sanitisers, common JavaScript functionalities like \texttt{document.write()} 
and \texttt{innerHTML} apply an \emph{implicit browser transduction} --- which
decodes HTML codes (e.g. \verb+&#39;+ is replaced by \verb+'+) in the input 
string --- before inserting the input string into the DOM.
Both of these examples can be expressed by (perhaps multiple) 
applications of the string-replace function.
Moreover, although these examples replace constants by constants, the popularity of template systems such as Mustache~\cite{Mustache} and Closure Templates~\cite{Closure} demonstrate the need for replacements involving variables.
Using Mustache, a web-developer, for example, may define an HTML fragment with placeholders that is instantiated with user data during the construction of the delivered page.

A regular expression (shortened as regex) is a sequence of characters that define a search pattern. Usually such patterns are used by string-searching algorithms for "find" or "find and replace" operations on strings, or for input validation.  


In programming language, stress regex, which admits features such as capturing groups, reference


symbolic execution requires SMT solving 

the main contribution is to introduce  a new transducer model, prioritized streaming string transducer (PSST) with the  motivation 
(1) using the priority to model regular expressions features such as capturing groups and (2) streaming transducers as a generic and expressive model for string manipulation. For instance, it can capture, among others, the reverse operation. 

The new contribution: 

\begin{itemize}
	\item extending Streaming String Transducers: \cite{FR17}
	
	\item  universal solution 
	
	\item 
\end{itemize} 
