
\documentclass{llncs}

%%%(20 pages max, must be anonymized)
%% Conference information

\usepackage{mathrsfs}
\usepackage{blindtext}
\usepackage{minted}
\usepackage{eqnarray}
\usepackage{amssymb}
\usepackage{mathtools}

\newif\ifdraft\drafttrue
%\newif\ifdraft\draftfalse
\ifdraft
\newcommand{\anthony}[1]{\color{red} {AL: #1 :LA} \color{black}}
\newcommand{\zhilin}[1]{\color{brown} {ZL: #1 :LZ} \color{black}}
\newcommand{\tl}[1]{\color{blue} {TL: #1 :LT} \color{black}}
\newcommand{\mat}[1]{\color{cyan} {MH: #1 :HM} \color{black}}
\newcommand{\philipp}[1]{\color{magenta} {PR: #1 :PR} \color{black}}
\newcommand{\zhilei}[1]{\color{violet} {ZLH: #1 :HZL} \color{black}}
\else
\newcommand{\anthony}[1]{}
\newcommand{\zhilin}[1]{}
\newcommand{\tl}[1]{}
\newcommand{\mat}[1]{}
\newcommand{\zhilei}[1]{}
\fi


%\newcommand{\ASSERT}[1]{\textsf{assert}\left(#1\right)}
%%%%%%%%%% End TeXmacs macros



\input{macros.tex}


\begin{document}

%% Title information
\title{Solving String Constraints With \\
Real-world Regular Expressions}



%% Author with single affiliation.
\author{}
\institute{}

\maketitle


%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%% some background on regular expressions
Regular expressions (RE) are a classical concept in formal language theory, which are expressions built from characters by the operators of concatenation, union, and repetition. Real-world regular expressions (RWRE) in programming languages differ from REs in the non-standard semantics of operators, e.g. non-commutative union and greedy repetition, as well as  the additional features like capturing groups and back references. 
%% symbolic execution requires faithful encoding of regex semantics
%String constraint solvers are one of the cornerstones of the analysis and verification of string manipulating programs. Faithful encoding of the semantics of real-world regular expressions in string constraint solvers facilitates more precise program analysis and verification.
%% state of the art string constraint solvers
%The semantics of real-word regular expressions are tricky and vary in different programming languages. It is challenging for string constraint solvers to support real-world regular expressions.
While REs are supported by every state-of-the-art string constraint solvers (e.g. CVC4 and Z3-str), none of them supports RWREs. The mismatch between REs in string constraint solvers and RWREs in programming languages hinders the precision and efficiency of symbolic executors, which have to resort to counter example guided abstraction and refinements (CEGAR) to mitigate this issue. 
%
%The semantics of real-world regular expressions are better to be encoded as faithfully as possible. For instance, in dynamic symbolic execution of string manipulating programs, in order to generate the inputs for execution paths and improve coverage,  the semantics of real-world regular expressions are better to be encoded as faithfully as possible in string constraint solvers. 
%% contribution
In this work, we propose an approach of natively supporting RWREs in string constraint solving. The key of our approach is to introduce a new automata model called prioritized streaming string transducers (PSST), 
which combines priorities in prioritized finite transducers and string variables in streaming string transducers, to model the string functions involving RWREs. 
%non-standard semantics of regular expression operators can be modeled by priorities and new features of capturing groups and back references can be modeled by string variables. 
%the non-standard semantics of regular expression operators can be modeled by priorities and new features of capturing groups and back references can be modeled by string variables. 
Based on PSSTs, we design a decision procedure for string constraints with RWREs. 
%% implementation and experiments
Moreover, we implement the decision procedure and do extensive experiments to evaluate its performance. The experimental results on over 160,000 string constraints generated from RWREs in open-source programs show that our approach dramatically improves the CEGAR-based approach for RWREs in both precision and efficiency. To the best of our knowledge, this work represents the first string constraint solver supporting RWREs.
\end{abstract}


%% Keywords
%% comma separated list
%\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission


%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.

\zhilin{
page limit: 20
}


%%%%%%%%%%%%%%%%%%%%%%%%%

\input{intro.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%

\input{motivating-example-new.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

%\input{prel.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

\input{rwre.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

\input{logic.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{psst-epsilon.tex}
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{decproc-new.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{impl-exp.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\input{experiments.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{alpha}
\bibliography{string}

\newpage
%% Appendix
\appendix

\input{appendix.tex}

\end{document}
