%!TEX root = main.tex


%This paper concerns strings and integers, two fundamental data-types in virtually all programming languages.
String-manipulating programs are notoriously subtle, and their potential bugs %Many subtle programming bugs are caused by , some with
may bring severe security consequences. A typical example is cross-site scripting
(XSS), which is among the OWASP Top 10 Application Security Risks~\cite{owasp17}.
Integer data type occurs naturally and extensively in string-manipulating programs. %For instance, access to lengths or positions of strings is frequently needed. 
%
An effective and increasingly popular method for identifying bugs, including XSS, is symbolic execution~\cite{symbex-survey}.
% and combinations with dynamic analysis
%called \emph{dynamic symbolic execution} \cite{jalangi,DART,EXE,CUTE,KLEE}.
%(See~\cite{symbex-survey} for an excellent survey.)
%Unlike purely random testing,
%which runs only \emph{concrete} program executions on different
%inputs, the techniques of symbolic execution 
In a nutshell, this technique analyses static paths
 through the program being considered.
Such a path can be viewed as a constraint~$\varphi$ over
appropriate data domains, and symbolic execution tools
demand fast constraint
solvers to check the satisfiability of $\varphi$. Such constraint
solvers need to support all native data-type operations occurring in
a program.
% (i.e., to check
%the \emph{feasibility} of the static path), which can be used for, e.g.,  generating
%inputs that lead to certain parts of the program or an erroneous behavior.

Typically, mainstream programming languages provide, apart from standard string functions (e.g., concatenation, $\replace$ and $\replaceall$), %string functions involving the integer data type, e.g. 
functions such as $\length$, $\indexof$, and $\substring$, which can convert strings to integers and vice versa. %, are also widely used in string-manipulating programs. 
These functions are indeed heavily utilised in practice; for instance, it was reported~\cite{Berkeley-JavaScript} that $\length$, $\indexof$, $\substring$, and variants thereof, comprise over 80\% of string function occurrences in 18 popular JavaScript applications, notably outnumbering concatenation. The introduction of integers exacerbates the intricacy of string-manipulating programs, and poses new theoretical and practical challenges in solver development. 


%in programs it , typically in the form of lengths or positions of strings. Besides the classical string functions e.g. concatenation $\concat$ and $\replaceall$, string functions involving the integer data type, e.g. $\length$, $\indexof$, and $\substring$, are also widely used in string-manipulating programs. For instance, it was reported in \cite{Berkeley-JavaScript} that $\length$, $\indexof$, $\substring$, and their variants, occupy more than 80 percent of string function occurrences in their subject Javascript applications (even much more frequently used than concatenation). This motivates \emph{string constraint solvers that reason about strings and integers simultaneously}. 



%$\length: \Sigma^* \rightarrow \Int$ (where $\Sigma$ is the alphabet and $\Sigma^*$ is the set of strings over $\Sigma$), $\substring: \Sigma^* \times \Int \times \Int \rightarrow \Sigma^*$, and $\indexof: \Sigma^* \times \Int \rightarrow \Int$.

%Reasoning about strings and integers is important and challenging, but is not well-studied. 
%It is easy to end up with undecidability: even the slightest extension of this theory (e.g. the theory with concatenation and letter counting)  would render the theory undecidable \cite{buchi,GB16}. 

When combining strings and integers, decidability can easily be lost; 
 for instance, the string theory with concatenation and letter counting functions is undecidable~\cite{buchi}.
\OMIT{
Although a great deal of research has shown that it is viable to reason about rather complex string operations without breaking decidability by %imposing proper 
restricting the form of constraints (e.g.,~\cite{CCH+18,CHL+19}; cf.\ related work for a brief survey), adding length constraints would immediately lead to undecidability~\cite{CCH+18}. 
}
It is still a major open problem whether the string theory with concatenation (arguably the simplest string operation) and length function 
(arguably the most common string-number function) is 
decidable~\cite{Vijay-length}. 
One promising approach to retain decidability is to enforce a syntactic
restriction to the constraints including \emph{solved forms}
\cite{Vijay-length}, acyclicity \cite{BFL13,Abdulla14,AbdullaA+19}, and 
straight-line fragment \cite{LB16,CCH+18,CHL+19,HJLRV18}. On the one hand,
such a restriction could lead to decidability of string constraint solving with 
complex string
constraints (not only concatenation, but also transducers), and integer
constraints (letter-counting, length constraints, IndexOf, etc.), e.g., see
\cite{LB16}. On the other hand, there is a lot of evidence (e.g. from 
benchmarking) that sufficiently many practical string constraints could satisfy 
such syntactic restrictions.
%
%At the same time, the classes of constraints that are required in practice sometimes demand solving techniques which are not necessarily complete. 
%
\OMIT{
Although there have been some recent results on decision procedures for string constraints involving the integer data-type, for instance~\cite{Vijay-length,LeH18,LinM18,LB16}, the theories considered in those approaches are usually quite restricted and do not reflect the constraints that occur in applications well. Overall, much more research is needed to
better understand the decidability of theories combining strings and integers.
}

On a more practical level, this leads to a situation where the vast majority of  state-of-the-art string constraint solvers, e.g., {\cvc}~\cite{cvc4}, {\zthree}~\cite{Z3-str3}, Trau~\cite{Abdulla17} (or its variants {\trauplus}~\cite{AbdullaA+19} and {\zthreetrau}~\cite{Z3-trau}),
%ABC \cite{ABC}, and SLENT \cite{WC+18}, 
resort to heuristics without completeness guarantees to support
reasoning about the combination of strings and integers. Such solvers
often show excellent performance in applications and on existing
benchmark suites, but heuristics tend to be brittle and difficult to
predict; for instance, sometimes innocent syntactic modifications of
constraints can cause solvers to diverge and not to terminate. There is still a gap between the string operations supported by
solvers, and the operations occurring in applications; for instance,
operations like \textsf{replaceAll} and escape/unescape-transformations
occur extremely frequently in programs, but are currently not handled well
by many solvers. 
The \emph{grand challenge} in string constraint solving is to
identify fragments and develop algorithms, that cover the constraints
occurring in \emph{applications}, while preserving \emph{decidability} and
admitting \emph{efficient implementations.}


%Nevertheless, solving string constraints involving the integer data type is far from trivial and actually undecidable in general \cite{buchi,CCH+18}. Therefore, all the aforementioned string constraint solvers
%They however inevitably  resort to heuristics without completeness guarantees. 

%Nevertheless, as already argued in~\cite{CHL+19}, despite the excellent performance of some of these solvers on some existing benchmark suites, decision procedures for string constraints with stronger theoretical guarantees, e.g., in the form of decidability (perhaps accompanied by a complexity analysis), are still highly desirable. 

%In a nutshell, there are at least two reasons: (1) they are \emph{theoretically intriguing} as this is essentially to delimit the (un)decidability boundary for the first-order theory of strings not least; and   
%(2) they are \emph{practically appealing} since they could potentially strike a good balance between precision and scalability, if they could be made an efficient implementation. 

%are amenable to implementation. 
%they can provide a kind of robustness guarantees upon which a string constraint solver could further improve and optimise.if they are amenable to implementation, 

%\zhilin{this paragraph is copied from the popl paper} Despite the excellent performance of some of these solvers on several existing benchmarks, there are good reasons for designing decision procedures with stronger theoretical guarantees, e.g., in the form of decidability (perhaps accompanied by a complexity analysis). One such reason is that string constraint solving is a research area in its infancy with an insufficient range of benchmarking examples to convince us that if a string solver works well on existing benchmarks, it will also work well on future benchmarks. A theoretical result provides a kind of robustness guarantee upon which a practical solver could further improve and optimise.

%(concatenation, length and regular constraints) 
%We tackle this issue 

%In contrast, in this paper we aim to develop decision procedures for string constraints involving the integer data type, which %, on the one hand, 
%\emph{provide completeness guarantees} and %on the other hand, 
%\emph{admit efficient implementation}.

\paragraph*{Contribution.} The main contribution of this paper is a decision procedure for an expressive class of string constraints involving the integer data-type, which includes not only concatenation, $\replaceall$, $\reverse$, finite transducers, and regular constraints, but also $\length$, $\indexof$ and $\substring$. The decision procedure is automata-theoretic and utilises a variant of cost-register automata introduced by Alur et al. \cite{RLJ+13}, which we call \emph{cost-enriched finite automata} (CEFA). The crux of the decision procedure is to compute the backward images of CEFAs under string functions,  in the same flavour as \cite{CHL+19} for string constraints \emph{without} the integer data type. Such an approach %inherits its elegance by 
is able to treat %the aforementioned seemly 
a wide range of string functions in a generic, and yet simple, way. To the best of our knowledge, the class of string constraints considered in this paper is currently the most expressive string theory involving the integer data type known to enjoy a decision procedure.
%\emph{the first decision procedure for string constraints that supports the string functions} $\substring$, $\length$, and $\indexof$, as well as concatenation, $\replaceall$, and finite transducers. \zhilin{improve latter}

Perhaps more importantly, our decision procedure admits an efficient implementation {\ostrich}+, which is based on %an extension of 
the recent {\ostrich} solver \cite{CHL+19}.  We have performed experiments on a wide range of benchmark suites, including the well-known {\kaluzabench} and {\pyexbench}, to evaluate the performance of {\ostrich}+. The results show that  %{\ostrich}+ achieves a nice balance between expressiveness and efficiency in the sense that 
1) {\ostrich}+ so far is the only string constraint solver capable of dealing with finite transducers and integer constraints, and 2) its performance is comparable with the best state-of-the-art string constraint solvers (e.g. {\cvc} and {\zthreetrau}) which are short of completeness guarantees. 

\input{related}

%%%%%%%%%%%%%%%%%%%%%%% The running example %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% The running example %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% The running example %%%%%%%%%%%%%%%%%%%%%%%
\input{runexmp.tex}
%%%%%%%%%%%%%%%%%%%%%%% The running example %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% The running example %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% The running example %%%%%%%%%%%%%%%%%%%%%%%
