 \documentclass{llncs}
\usepackage{minted}
%\documentclass[envcountsame, fleqn]{llncs}
  \usepackage{amsmath, amssymb, latexsym}
  \usepackage{graphicx}
\usepackage{epic,eepic}

%\pagestyle{plain}
\usepackage{listings}
\usepackage{psfrag}
\usepackage{rotating}

\usepackage{url}
\usepackage{amssymb,amsthm, epsfig,amstext}
\usepackage{txfonts}
\usepackage{algorithmic}
\usepackage[linesnumbered,ruled,noend]{algorithm2e}
\usepackage{graphicx}

\usepackage{mathrsfs}

\usepackage{color, colortbl}




\setminted{numbersep=5pt, xleftmargin=12pt}
\usemintedstyle{friendly} 

\usepackage{todonotes}
\usepackage{multirow}
\usepackage{float,color}
\usepackage{picinpar,color,xcolor,wrapfig}
\setcounter{secnumdepth}{4}
\sloppy

\pagestyle{plain}

%%%%%%%%%%%%%%%   MACROS  %%%%%%%%%%%%%%%%%%


\input{macros.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\zhilin}[1]{\color{cyan} {ZL: #1 :LZ} \color{black}}
\newcommand{\tl}[1]{\color{purple} {TL: #1 :LT} \color{black}}
\newcommand{\anthony}[1]{\color{blue} {AL: #1 :LA} \color{black}}
\newcommand{\matt}[1]{\color{blue} {MH: #1 :HM} \color{black}}
\newcommand{\philipp}[1]{\color{blue} {PR: #1 :RP} \color{black}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\title{A Decision Procedure for Path Feasibility of \\
String-Manipulating Programs\\
with Integer Data Type}


\author{}

\institute{ }

\begin{document}


%Regular Papers should not exceed 20 pages in LNCS format, not counting references and appendices.
%
%Regular papers at CAV 2020 will follow a full double blind review process, which means that author names and affiliations must be omitted from the submission. Additionally, if a submission refers to prior work done by the authors, the reference should be made in the third person. These are firm submission requirements, and any regular paper that does not conform to these requirements will be rejected without review.
%
%Authors can include a clearly marked appendix at the end of their submissions that is exempt from the page limit restrictions. However, the reviewers are not obliged to read the contents of these appendices. These papers should contain original research and sufficient detail to assess the merits and relevance of the contribution. Papers will be evaluated on basis of a combination of correctness, technical depth, significance, novelty, clarity, and elegance.

\maketitle

\begin{abstract}
Strings are widely used in programs, especially in Web programs, e.g. JavaScript and PHP. The integer data type naturally occurs in string-manipulating programs, typically being used to refer to length of, or position in, strings. 
%For instance, the inputs or outputs of widely used string operations including length, substring, and indexof involve the integer data type.
Analysis and testing of string-manipulating programs can be formulated as the path feasibility problem by symbolic execution, which is to decide whether a given execution path is feasible, namely, whether there exist assignments to the input variables so that a given path can execute to the end. %, which plays a central role in the static analysis and verification, e.g. symbolic execution, of programs. 
%Path feasibility problem of string-manipulating programs is very challenging and undecidable in general, especially when involving the integer data type. 
%Aiming at solving the path feasibility of string-manipulating programs in practice, 
%on the one hand, most of the 
State-of-the-art string constraint solvers usually provide strong support of complex string operations, in conjunction with limited support of integer data type. However, partially because this is a generally undecidable problem, %, and on the other hand, 
they mainly resort to heuristics without completeness guarantees. \\
In this paper, we propose a decision procedure of the path feasibility problem for a class of string-manipulating programs
which not only support complex string operations %involving only the string data type, e.g. 
such as concatenation, replaceAll, reverse, and finite state transducers, but also operations involving both the string and integer data type, e.g., length, substring, and indexof. To the best of our knowledge, this represents the most expressive string constraint language which is currently known to be decidable. 
Our decision procedure is automata-theoretic which is based on a variant of the cost register automata (Alur et al. LICS 2013). 
%To the best of our knowledge, %this is %the first time that 
%the first (complete) decision procedure which covers the most expressive string constraint language %have been achieved for the path feasibility problem of such an expressive class of string-manipulating programs. Furthermore, 
Different from most string constraint solving techniques which provide completeness guarantees, 
our decision procedure is amenable to implementation. 
%We implement the decision procedure on top of OSTRICH, resulting into a new solver called 
The implementation gives rise to OSTRICH+, which  is based on a recent solver OSTRICH which could not tackle integer data types. 
We utilise a wide range of benchmarks to evaluate the performance of OSTRICH+. The experimental results show that OSTRICH+ is competitive with some of the best state-of-the-art string constraint solvers.
\end{abstract}

%\section{Introduction}
%\label{intro}
%
%\section{Preliminaries}
%\label{prel}
%
%Definition for NFA, NFT.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%20 pages, excluding references.
%
%Abstract+Conclusion: 1
%
%Introduction: 3
%
%Prelimilary: 2
%
%Logic: 2
%
%Decision procedure: 5-6
%
%Implementation: 3
%
%Evaluation:3


\input{intro.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}\label{sec:prel}

\input{prel.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{String-Manipulating Programs with Integer Data Type}\label{sec:logic}

\input{logic.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Automata-Theoretic Foundations}\label{sec:cefa}

%\input{cefa.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Decision Procedures for Path Feasibility}\label{sec:dec}

\input{decision.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Implementation}

\input{impl.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Evaluation}

\input{eval.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage 
\bibliographystyle{abbrv}
\bibliography{string}

%\iffalse
\newpage
%\setcounter{page}{1}

\begin{appendix}
\input{appendix}
\end{appendix}

%\fi

\end{document}
