%!TEX root = main.tex

\section{Implementation and Experiments}
\label{sect:impl}

%For our experiments, we have implemented an SMT solver,
%\ostrich, %\footnote{Name anonymized for doubly-blind review, and will  be provided in the final version.} 
%based on the calculus for
%$\strline$.  \ostrich\ extends the open-source solver
%OSTRICH~\cite{CHL+19}, and is able to decide satisfiability of
%$\strlinesl$ formulas. \ostrich\ also inherits support for most of the
%other operations of the SMT-LIB theory of Unicode
%strings\footnote{\url{http://smtlib.cs.uiowa.edu/theories-UnicodeStrings.shtml}}
%from OSTRICH.

For our experiments, we have extended the open-source solver
OSTRICH~\cite{CHL+19} to support for $\strline$ %implemented %an SMT solver,
%\ostrich,  
based on the calculus. In particular, it is able to decide the satisfiability of
$\strlinesl$ formulas. 
The new extension can handle most of the
other operations of the SMT-LIB theory of Unicode
strings.\footnote{\url{http://smtlib.cs.uiowa.edu/theories-UnicodeStrings.shtml}}
%from OSTRICH.

%\cite{CHL+19}, %\url{https://github.com/uuverifiers/ostrich}}
%which provides a modular and easy-to-use framework for extending all
%sorts of string operations. 
%As shown in Section \ref{sec:decision},
%PSSTs satisfy the conditions required by the backward reasoning
%approach of OSTRICH, which enables us to integrate our logic with
%standard string theory. 


\subsection{Implementation}

Our solver extends classical regular expressions in SMT-LIB
with indexed {\sf re.capture} and {\sf re.reference} operators, which
denote capturing groups and references to them. We also add {\sf re.*?}, {\sf re.+?}, {\sf re.opt?} and {\sf re.loop?} as the lazy counterparts of
Kleene star, plus operator, optional operator and loop operator.

Three new string operators are introduced to make use of these extended regular
expressions: {\sf str.replace\_cg}, {\sf str.replace\_cg\_all}, and
{\sf str.extract}. The operators {\sf str.replace\_cg} and {\sf
  str.replace\_cg\_all} are the counterparts of the standard {\sf
  str.replace\_re} and {\sf replace\_re\_all} operators, and allow
capturing groups in the match pattern and references in the
replacement pattern. As an example, the following constraint swaps the
first name and the last name, as in Example~\ref{exmp-name-swap}:
%lower-case characters~$x$ with a sub-sequent upper-case character~$Y$:

{\small
\begin{minted}{lisp}
(= w (str.replace_cg_all v
      (re.++ ((_ re.capture 1) (re.+ (re.union (re.range "A" "Z") (re.range "a" "z"))))
             (str.to.re " ") 
             ((_ re.capture 2) (re.+ (re.union (re.range "A" "Z") (re.range "a" "z")))))
      (re.++ (_ re.reference 2) (_ re.reference 1))))
\end{minted}
}

The replacement string is written as a regular expression only
containing the operators {\sf re.++}, {\sf str.to\_re}, and {\sf
  re.reference}. The use of string variables in the replacement
parameter is not allowed, since the resulting transformation could
not be mapped to a \PSST.

The indexed operator {\sf str.extract} implements $\extract_{i, e}$ in
$\strline$. For instance,

{\small
\begin{minted}{lisp}
  ((_ str.extract 1)
      (re.++ (re.*? re.allchar) ((_ re.capture 1) (re.+ (re.range "a" "z")) re.all)) x)
\end{minted}
}

\noindent
extracts the left-most, longest sub-string consisting of only lower-case
characters from a string~$x$.

Our implementation is able to handle \textit{anchors} as well,
although for reasons of presentation we did not introduce them as part
of our formalism. Anchors match certain
positions of a string without consuming any input characters. In most
programming languages, it is common to use \verb!^!
and
\verb!$! in regular expressions to signify the start and end of a
string, respectively. We add \textsf{re.begin-anchor} and
\textsf{re.end-anchor} for them. Our implementation correctly models
the semantics of anchors and is able to solve constraints containing
these operators.

\ostrich\ implements the procedure in Theorem~\ref{thm-main}, and
focuses on SL formulas. The three string operators mentioned above
will be converted into an equivalent PSST (see \ifproceeding the full
version of the paper \cite{popl22-full}).  \else
Appendix~\ref{appendix:sec-extract-replace-to-psst}). \fi {\ostrich}
then iteratively applies the propagation rules from
Section~\ref{sect:calculus} to derive further \regexp\ constraints,
and eventually either detect a contradiction, or converge and find a
fixed-point. For straight-line formulas, the existence of a
fixed-point implies satisfiability, and a solution can then be
constructed as described in Section~\ref{sect:calculus}.  In addition,
similar to other SMT solvers, {\ostrich} applies simplification rules
(e.g., {\sf Fwd-Prop-Elim, =-Prop, Subsume, Close}, etc in
Table~\ref{tab:calculus}) to formulas before invoking the SL
procedure.  This enables \ostrich\ to solve some formulas outside of the SL
fragment, but is not a complete procedure for non-SL formulas.

\begin{figure}[tb]
  \scriptsize

  \begin{minipage}{0.55\linewidth}
\begin{minted}{lisp}
(declare-fun x () String)
(define-fun  y () String (str.replace_cg_all x <re1> <repl>))
(push 1)
(assert (str.in.re x (re.++ re.all <re1> re.all)))
(assert (str.in.re y (re.++ re.all <re2> re.all)))
(check-sat) (get-model)
(pop 1) (push 1)
(assert (str.in.re x (re.++ re.all <re1> re.all)))
(assert (not (str.in.re y 
          (re.++ re.all <re2> re.all))))
(check-sat) (get-model)
(pop 1) (push 1)
(assert (not (str.in.re x (re.++ re.all <re1> re.all))))
(check-sat) (get-model)
(pop 1)
\end{minted}
  \end{minipage}\hfill
  \raisebox{-19.2ex}{\rule{0.4pt}{40ex}}\hfill
  \begin{minipage}{0.43\linewidth}
\begin{minted}{js}
function fun(x) {
   if(/<re1>/.test(x)) {
      var y = x.replace(/<re1>/g, <repl>);
      if(/<re2>/.test(y))
        console.log("1");
      else
        console.log("2");
   }
   else
       console.log("3");
}

var S$ = require("S$");
var x = S$.symbol("x", "");
fun(x);
\end{minted}
  \end{minipage}
  
  \caption{Harnesses with replace-all: SMT-LIB for \ostrich\ (left),
    and JavaScript for \expose{} (right).}
  \label{fig:harness}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Experimental evaluation}

Our experiments have the purpose of answering the following main questions:

%\smallskip
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\OMIT
{
\textbf{R1:} Are the  {\regexp}s defined in this paper
suitable to encode regular expressions in programming languages,
for instance ECMAScript regular expressions~\cite{ECMAScript11}?\zhilin{Should R1 be removed, since we already have semantics validation experiments ?}\philipp{yes, I think we should remove it; everybody agrees?}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent
\textbf{R1:} How does \ostrich\ compare to other solvers that can
handle real-world regular expressions, including greedy/lazy
quantifiers and capturing groups?
\\
\textbf{R2:} How does \ostrich\ perform in the context of symbolic execution,
the primary application of string constraint solving?

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\OMIT
{
\paragraph{For \textbf{R1}:} We implemented a translator from ECMAScript 11th Edition (ES11 for short) regular
expressions to \regexps, and integrated \ostrich\ into the symbolic
execution tool Aratha~\cite{aratha}. We then ran Aratha+\ostrich\ on
the regression test suite of \expose{}~\cite{DBLP:conf/spin/LoringMK17},
as well as some other JavaScript programs containing match or replace
functions extracted from Github. To verify the soundness of
Aratha+\ostrich, we compared the results with those produced by
\expose{}; we also checked the correctness of models computed by
\ostrich\ by concretely executing the JavaScript program under test on
the generated inputs, to confirm that the concrete execution indeed
follows the targeted path. The results are summarized in Table~\ref{tab:exp-r1};
no inconsistencies were observed in the experiments, showing that the
semantics in this paper are indeed suitable for capturing ES11
semantics.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\smallskip
{\em For \textbf{R1}:} There are no standard string benchmarks
involving \regexps, and we are not aware of other constraint solvers
supporting capturing groups, neither among the SMT nor the CP
solvers. %(e.g.,
%\cite{DBLP:conf/cpaior/ScottFPS17,DBLP:conf/cp/AmadiniGS20}).
The closest related work is the algorithm implemented in \expose{}, which
applies Z3~\cite{Z3} for solving string constraints, but augments
it with a refinement loop to approximate the {\regexp}
semantics.\footnote{We considered replacing Z3 with \ostrich\ in
  \expose{} for the experiments. However, \expose{} integrates Z3 using its
  C API, and changing to \ostrich, with native support
  for capturing groups, would have required the rewrite of substantial
  parts of \expose{}.}
%
For \textbf{R1}, we compared \ostrich\ with \expose{}+Z3 on 98,117
\regexps\ taken from \cite{DMC+19}.

For each of the regular expressions, we created four harnesses: two in
SMT-LIB, as inputs for \ostrich, and two in JavaScript, as inputs for
\expose{}+Z3. The two harnesses shown in Fig.~\ref{fig:harness} use one of the
regular expressions from \cite{DMC+19} (\verb!<re1>!) in combination with
the replace-all function to simulate typical string processing;
\verb!<re2>! is the fixed pattern \verb![a-z]+!, and \verb!<repl>! the
replacement string \verb!"$1"!. The three paths of the JavaScript
function~\verb!fun! correspond to the three queries in the SMT-LIB
script, so that a direct comparison can be made between the results of
the SMT-LIB queries and the set of paths covered by \expose{}+Z3. The other
two harnesses are similar to the ones in Fig.~\ref{fig:harness}, but
use the match function instead of replace-all, and contain four
queries and four paths, respectively.

\begin{table}[t]

  \begin{center}
  \begin{tabular}{|l@{~~}|*{6}{c}|*{5}{c}@{~~}|}
    \hline
     & 
    \multicolumn{6}{c|}{\textbf{\ostrich}} &
    \multicolumn{5}{c|}{\textbf{\expose{}+Z3}}
    \\
      & \multicolumn{6}{c|}{\# queries solved within 60s}
      & \multicolumn{5}{c|}{\# paths covered within 60s}
    \\
     & 0 & 1 & 2 & 3 & 4 & \#Err
     & 0 & 1 & 2 & 3 & 4
    \\\hline
    \textbf{Match}  & 422 & 249 & 751 & 386 & 95,175 & 1,134
    & ~3,333 & 9,274 & 36,916 & 48,594 & 0
    \\
     \emph{(98,117} & \multicolumn{6}{c|}{Average time: 1.57s}
    &\multicolumn{5}{c|}{Average time: 28.0s}
    \\
    \emph{~~benchm.)} & \multicolumn{6}{c|}{Total \#sat: 250,947, \#unsat: 132,662}
    & \multicolumn{5}{c|}{Total \#paths covered: 228,888}
    \\\hline
    \textbf{Replace} & 4,170 & 2,463 & 555 & 89,794 & --- & 1,135
    & ~5,281 & 18,221 & 69,059 & 5,556 & ---
    \\
    \emph{(98,117} & \multicolumn{6}{c|}{Average time: 6.62s}
    & \multicolumn{5}{c|}{Average time: 55.0s}
    \\
    \emph{~~bench.)} & \multicolumn{6}{c|}{Total \#sat: 259,354, \#unsat: 13,601}
    & \multicolumn{5}{c|}{Total \#paths covered: 173,007}
      \\\hline
  \end{tabular}
  \end{center}
  \caption{The number of queries answered by \ostrich, and number of
    paths covered by \expose{}+Z3, in \textbf{R1}. 
    Experiments were done on an AMD Opteron 2220 SE machine, running
    64-bit Linux and Java~1.8.  Runtime per benchmark was limited to
    60s wall-clock time, 2GB memory, and the number of tests
    executed concurrently by \expose{}+Z3 to 1.  Average time is
    wall-clock time per benchmark, timeouts count as 60s.}
  \label{tab:exp-r2}

	\begin{center}
	\begin{tabular}{|l@{\quad}|*{6}{c}|*{6}{c}|}
	\hline
	   & 
	  \multicolumn{6}{c|}{\textbf{Aratha+\ostrich}} &\multicolumn{6}{c|}{\textbf{\expose{}+Z3}}
	  \\
    & 
	  \multicolumn{6}{c|}{~\# paths covered within 60s~~} &\multicolumn{6}{c@{\quad}|}{~\# paths covered within 60s}
	  \\
	   & ~~0~  & ~1~ &  ~2~ & ~3~ & ~$\geq$4~ & ~\#T.O.~ &
	    ~~0~  & ~1~ &  ~2~ & ~3~ & ~$\geq$4~ & ~\#T.O.~
	  \\\hline
	  \textbf{\expose{}} &  14 & 9 & 9 & 2 & 15 & 0  & 14 &  9 & 9 & 2 & 15 & 6  
	  \\
	  \emph{(49 programs)} & \multicolumn{6}{c|}{Average time: \textbf{4.66}s} & \multicolumn{6}{c@{\quad}|}{Average time: 57.44s}\\
	  & \multicolumn{6}{c|}{Total \#paths covered:124}  & \multicolumn{6}{c@{\quad}|}{Total \#paths covered:121}	  
	  \\\hline
	  \textbf{Match} & 3 & 7 & 12 & 6 & 0 & 0  & 3 & 8 & 12 & 5 & 0 & 6
	  \\
	  \emph{(28 programs)} & \multicolumn{6}{c|}{Average time: \textbf{5.19}s} & \multicolumn{6}{c@{\quad}|}{Average time: 60.26s}\\
	  & \multicolumn{6}{c|}{Total \#paths covered: 49}  & \multicolumn{6}{c@{\quad}|}{Total \#paths covered: 47}	  
	  \\\hline
	  \textbf{Replace} & 12 & 20 & 6 & 0 & 0 & 0  & 15 & 21 & 2 & 0 & 0 & 23
	  \\
	  \emph{(38 programs)} & \multicolumn{6}{c|}{Average time: \textbf{4.14}s} & \multicolumn{6}{c@{\quad}|}{Average time: 95.34s}\\
	  & \multicolumn{6}{c|}{Total \#paths covered: 32}  & \multicolumn{6}{c@{\quad}|}{Total \#paths covered: 25}	  
	  \\\hline
	\end{tabular}
	\end{center}
	\caption{Results of Expose+Z3 and Aratha+{\ostrich} on Javascript programs for \textbf{R2}. Experiments were done on an Intel(R)-Core(TM)-i5-8265U-CPU-@1.60GHz cpu, running 64-bit Linux and Java 1.8. Runtime was limited to 120s wall-clock time. Average time is
    wall-clock time needed per benchmark, and counts timeouts as 60s. \#T.O.\& Err is the number of timeouts. Note that some paths may have already been covered before T.O. }
	\label{tab:exp-r1}
\end{table}



The results of this experiment are shown in
Table~\ref{tab:exp-r2}. \ostrich\ is able to answer all four queries
in 95,175 of the match benchmarks (97\%), and all three queries in
89,794 of the replace-all benchmarks (91.5\%). The errors in 1,134
cases (resp., 1,135 cases) are mainly due to back-references in
\verb!<re1>!, which are not handled by \ostrich. \expose{}+Z3 can
cover 228,888 paths of the match problems in total (91.2\% of the
number of sat results of \ostrich), although the runtime of
\expose{}+Z3 is on average 18x higher than that of \ostrich. For
replace, \expose{}+Z3 can cover 173,007 paths (66.7\%), showing that
this class of constraints is harder; the runtime of \expose{}+Z3 is on
average 8x higher than that of \ostrich.  Overall, even taking into
account that \expose{}+Z3 has to analyze JavaScript code, as opposed
to the SMT-LIB given to \ostrich, the experiments show that \ostrich\
is a highly competitive solver for \regexps.

\smallskip
{\em For \textbf{R2}:} For this experiment,
we  integrated \ostrich\ into the symbolic
execution tool Aratha~\cite{aratha}.  We compare Aratha+{\ostrich} with
\expose{}+Z3 on the regression test suite of \expose{}~\cite{DBLP:conf/spin/LoringMK17},
as well as a collection of other JavaScript programs containing match or replace
functions extracted from Github. In
Table~\ref{tab:exp-r1}, we can see that Aratha+{\ostrich} can within
60s cover slightly more paths than \expose{}+Z3. Aratha+{\ostrich} can
discover feasible paths more quickly than \expose{}+Z3, however: on
all three families of benchmarks, Aratha+{\ostrich} terminates on
average in less than 10s, and it discovers all paths 
within 35s. \expose{}+Z3 needs the full 60s for
29 of the programs (``T.O.'' in the table),
and it finds new paths until the end of the 60s. Since \expose{}+Z3
handles the replace-all operation by unrolling, it is not able to
prove infeasibility of paths involving such operations, and will
therefore not terminate on some programs.
%is faster than \expose{}+Z3 by a factor between 1.7 and 15, while being
%able to cover more paths. The smaller speedup compared to the results
%for \textbf{R2} can be explained by the fact that Aratha should spend time in processing Javascript code, besides using {\ostrich} to solve path constraints.
%SMT-LIB queries
%produced by Aratha are not pure string constraints, but also include
%ADTs, arrays, and bit-vectors. 
Overall, the experiments indicate that {\ostrich} is more
efficient than the CEGAR-augmented symbolic execution for dealing
with \regexps.



