%!TEX root = popl2018.tex

\section{Related work}\label{sec-rel}

 
We now discuss some related work. We split our discussion into two categories: (1) theoretical results in terms of decidability and complexity; (2) practical (but generally incomplete) approaches used in string solvers.  We emphasise work on $\replaceall$ functions as they are our focus. 

\paragraph{Theoretical Results}
We have discussed in Section \ref{sec:intro} works on string constraints with 
the theory of strings with concatenation. This research programme builds on
the question of solving satisfiability of \emph{word equations}, i.e., a
string equation $\alpha = \beta$ containing concatenation of string constants
and variables. Makanin showed decidability \cite{Makanin}, whose upper bound
was improved to PSPACE in \cite{P04} using a word compression technique. 
A simpler algorithm was in recent years proposed in \cite{J17} using
the recompression technique. The best lower bound for this problem is still NP,
and closing this complexity gap is a long-standing open problem. Decidability
(in fact, the PSPACE upper bound) can be retained in the presence of regular
constraints (e.g. see \cite{Schulz}). This can be extended to existential theory
of concatenation with regular constraints using the technique of \cite{buchi}.
The replace-all operator cannot be expressed by the concatenation operator alone. For 
this reason, our decidability of the fragment of $\strline[\replaceall]$ cannot
be derived from the results from the theory of concatenation alone.

%Our work is closely related to solving word equations, which are a conjunction of equations of $v=w$, where $v, w$ are concatenation of string constants and variables. The computational complexity of this problem remains unknown, with the best lower and upper bounds being NP and PSPACE respectively. Makanin %refuted a conjecture 
%showed, perhaps surprisingly, that the problem is decidable \cite{Makanin}, and %Plandowski's  on the decidability and complexity of satisfiability for word equations, i.e., 
%Plandowski explored an approach of compression and proposed a PSPACE algorithm \cite{P04}.  This is the best bound up to date, though a simpler PSPACE algorithm with smaller space requirement (nondeterministic $O(n \log n)$ space) was proposed by Jez \cite{J16}, based on compression. Very recently, Jez \cite{J17} further improves the complexity to (nondeterministic) $O(n )$, i.e.,  linear space. 

%The connection between compression and word equations was first observed by Plandowski and Rytter \cite{PR98}. %, who showed that a length-minimal solution of size N has a compressed representation
%of size poly(n, log N). 
\OMIT{
Essentially, the constraint language  $\strline[\replaceall]$ studied in this paper is word equations where the $\replaceall$ functions, which subsume the concatenation operation, are used. However, the additional straight-line restrictions are imposed.  
}

Regarding the extension with length constraints, it is still a long-standing
open problem whether word equations with length constraints is decidable, though
it is known that letter-counting (e.g. counting the number of occurrences of 0s
and 1s separately) yields undecidability \cite{buchi}. It was shown in
\cite{LB16} that the length constraints (in fact, letter-counting) can be
added to the subclass of $\strline[\replaceall]$ where the pattern/replacement 
are constants, while preserving decidability. In contrast, if we allow 
variables on the replacement parameters of formulas in $\strline[\replaceall]$,
we can easily encode the Hilbert's 10th problem with length (integer) 
constraints. 
%In fact, this undecidability holds even if we use the unary
%alphabet $\Sigma = \{a\}$, and that the pattern string is
%always fixed to the letter $a$.

%Solving word equation was an intriguing problem since the beginning of computer science, investigated
%initially due to its ties to Hilbert’s 10th problem. Initially it was conjectured that this
%problem is undecidable, which was disproved by Makanin [10]. At the beginning little attention
%was given to computational complexity of Makanin’s algorithm and the problem itself; these questions
%were reinvestigated in the ’90 [6, 18, 9], culminating in the EXPSPACE implementation of
%Makanin’s algorithm by Gutiérrez [5].


%\cite{J17}
%Word equations in linear space
%
%Word equations are an important problem on the intersection of formal languages and algebra. Given two sequences consisting of letters and variables we are to decide whether there is a substitution for the variables that turns this equation into true equality of strings. The computational complexity of this problem remains unknown, with the best lower and upper bounds being NP and PSPACE. Recently, a novel technique of recompression was applied to this problem, simplifying the known proofs and lowering the space complexity to (nondeterministic) O(n log n). In this paper we show that word equations are in nondeterministic linear space, thus the language of satisfiable word equations is context-sensitive. We use the known recompression-based algorithm and additionally employ Huffman coding for letters. The proof, however, uses analysis of how the fragments of the equation depend on each other as well as a new strategy for nondeterministic choices of the algorithm, which uses several new ideas to limit the space occupied by the letters.


\OMIT{
%A Decision Procedure for String Logic with Equations, Regular Membership and Length Constraints 
 Le \cite{L16} considered the satisfiability problem for string logic with word equations, regular membership and Presburger constraints over length functions. %The difficulty comes from multiple occurrences of string variables making state-of-the-art algorithms non-terminating. Our main contribution is to 
He showed that the satisfiability problem in a fragment where no string variable occurs more than twice in an equation is decidable. 
%In particular, he proposed a semi-decision procedure for arbitrary string formulae with word equations, regular membership and length functions, and showed that the algorithm always terminates for the aforementioned decidable fragment, with a complexity analysis. 
This work is largely distant from ours, as $\replaceall$ was not addressed. However, we mention that the fragment considered by Le allows Presburger constraints over length functions.
%The essence of our procedure is an algorithm to enumerate an equivalent set of solvable disjuncts for the formula. We further show that the algorithm always terminates for the aforementioned decidable fragment. Finally, we provide a complexity analysis of our decision procedure to prove that it runs, in the worst case, in factorial time.

%In \cite{BTV09} the authors discussed the problem of path feasibility for programs manipulating strings using a collection of standard string library functions. They prove results on the complexity of this problem, including its undecidability in the general case and decidability of some special cases. \tl{how would it connect to ours?}
}



The $\replaceall$ function can be seen as a special, yet expressive, string transformation function, aka string transducer. From this viewpoint, the closest work is~\cite{LB16}, which we discuss extensively in the introduction. Here, we discuss two further 
%the $\replaceall$ function is also related to two 
recent transducer models: streaming string transducers \cite{AC10} and symbolic transducers \cite{symbolic-transducer}. 

A streaming string transducer is a finite state machine where  a finite set of string variables are used to store the intermediate results for output. The $\replaceall(x, e, y)$ term can be modelled by an extension of streaming string transducers \emph{with parameters}, that is, a streaming string transducer which reads an input string (interpreted as the value of $x$), uses $y$ as a free string variable which is presumed to be read-only, and updates a string variable $z$, which stores the computation result, by a string term which may involve $y$. Nevertheless, to the best of our knowledge, this extension of streaming string transducers has not been investigated so far. 

Symbolic transducers are an extension of Mealy machine to infinite alphabets by using a variable $cur$ to represent the symbol in the current position, and replacing the input and output letters in transitions with unary predicates $\varphi(cur)$ and terms involving $cur$ respectively. Symbolic transducers can model $\replaceall$ functions \emph{when the third parameter is a constant}. Inspired by symbolic transducers, it is perhaps an interesting future work to consider an extension of the $\replaceall$ function by allowing predicates as patterns. %to sequences of numerical values. 
For instance, one may consider the term $\replaceall(x, cur \equiv 0 \bmod 2, y)$ which replaces every even number in $x$ with $y$. %This, however, is not addressed in the current paper. 

Finally, the $\replaceall$ function is related to Array Folds Logic introduced by Daca et al \cite{DHK16}. The authors considered an extension of the quantifier-free theory of integer arrays with counting. The main feature of the logic is the \emph{fold} terms, borrowed from the folding concept in functional programming languages. Intuitively, a fold term applies a function to every element of the array to compute an output. If strings are treated as arrays over a finite domain (the alphabet), the $\replaceall$ function can be seen as a fold term. Nevertheless, the $\replaceall$ function goes beyond the fold terms considered in \cite{DHK16}, since it outputs a string (an array), instead of an integer. Therefore, the results in \cite{DHK16} cannot be applied to our setting.

\paragraph{Practical Solvers}
A large amount of recent work develops practical string solvers
including  Kaluza~\cite{Berkeley-JavaScript}, Hampi~\cite{HAMPI}, 
Z3-str~\cite{Z3-str}, CVC4~\cite{cvc4}, Stranger~\cite{YABI14}, Norn~\cite{Abdulla14}, S3 and S3P~\cite{S3,TCJ16}, and FAT~\cite{Abdulla17}.
Among them, only Stranger, S3, and S3P support $\replaceall$.  

%String solvers that support concatenations and the replace-all operator are available. \cite{BTV09,TCJ14,YABI14,S3}

%\cite{BTV09} considered an efficient finite model finding method for string constraints. They develop a two-tier finite model finding procedure. First, an integer abstraction of string constraints are passed to an SMT (Satisfiability Modulo Theories) solver. The abstraction is either unsatisfiable, or the solver produces a model that fixes lengths of enough strings to reduce the entire problem to be finite domain. The resulting fixed-length string constraints are then solved in a second phase. 

%We implemented the procedure in a symbolic execution framework, report on the encouraging results and discuss directions for improving the method further.


In the Stranger tool, %Verifying string manipulating programs is a crucial problem in computer security. String operations are used extensively within web applications to manipulate user input, and their erroneous use is the most common cause of security vulnerabilities in web applications. We 
an automata-based approach was provided for symbolic analysis of PHP programs, where two different semantics $\replaceall$ were considered, namely, the leftmost and longest matching as well as the leftmost and shortest matching. Nevertheless, they focused on the abstract-interpretation based analysis of PHP programs and provided an \emph{over-approximation} of all the possible values of the string variables at each program point. Therefore, their string constraint solving algorithm is \emph{not} an exact decision procedure. In contrast, we provided a decision procedure for the straight-line fragment with the rather general $\replaceall$ function, where the pattern parameter can be arbitrary regular expressions and the replacement parameter can be variables. In the latter case,  we consider the leftmost and longest semantics mainly for simplicity, and the decision procedure can be adapted to the leftmost and shortest semantics easily.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
 For the string operations, the authors focus on two common ones: concatenation and replacement. The latter is close to---but not the same as---the $\replaceall$ function considered in this paper. However, in this paper, Yu et al provided   an \emph{over-approximation} of more %restricted replace 
commonly used semantics, i.e., the longest match and first match semantics. 
%
%\cite{SMV12} translating regular expression matching into transducer. 
%
They use deterministic finite automata (DFAs) to represent possible values of string variables. Using forward reachability analysis we compute an over-approximation of all possible values that string variables can take at each program point. They also implemented Stranger, an automata-based string analysis tool, with experiments. In general, this is essentially an abstract interpretation based approach.  In comparison, our algorithm is also automata-based, but we work on a semantics of $\replaceall$, but not its approximation. \tl{more need to be said here}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Intersecting these with a given attack pattern yields the potential attack strings if the program is vulnerable. Based on the presented techniques, we have implemented Stranger, an automata-based string analysis tool for detecting string-related security vulnerabilities in PHP applications. We evaluated Stranger on several open-source Web applications including one with 350,000+ lines of code. Stranger is able to detect known/unknown vulnerabilities, and, after inserting proper sanitization routines, prove the absence of vulnerabilities with respect to given attack patterns.

 
The S3 and S3P tools also support the $\replaceall$ function, where some
progressive searching strategies were provided to deal with the non-termination
problem caused by the recursively defined string operations (of which 
$\replaceall$ is a special case). Nevertheless, the solvers are 
incomplete as reasoning about unbounded strings defined recursively is in 
general an undecidable problem.


%, the authors present S3 (which stands for Symbolic String Solver), a new symbolic string solver.
%The constraint language covers all the main string operations including the replace-all function. The authors 
%provided algorithms which make use of a symbolic representation so that membership in a set defined by a regular expression can be encoded as string equations. 

%To amplify this point, let us now state some statistics from a comprehensive
%study of practical JavaScript applications [28]. Constraints
%arising from the applications have an average (per benchmark
%query) of 63 JavaScript string operations, while the remaining
%are boolean, logical and arithmetic constraints. The largest fraction
%are for operations like indexOf, length (78%). A significant
%fraction of the operations, including substring (5%), replace
%(8%), and split, match (1%). Of the match, split and
%replace operations, 31% are based on regular expressions. Operations
%such as replace and split give rise to new strings
%from the original ones, thereby giving rise to constraints involving
%multiple string variables.



%. The algorithm first makes use of a symbolic representation
%so that membership in a set defined by a regular expression
%can be encoded as string equations. Secondly, there is a constraint based
%generation of instances from these symbolic expressions so
%that the total number of instances can be limited. 
%
%We evaluate S3 on a well-known set of practical benchmarks, demonstrating both
%its robustness (more definitive answers) and its efficiency (about 20
%times faster) against the state-of-the-art.



\hide{
%Progressive Reasoning over Recursively-Defined Strings
\cite{TCJ16} 
Trinh et al considered %the problem of reasoning over an expressive constraint language for unbounded strings. 
%In particular, they considered 
recursively defined string functions, a very expressive way to define functions manipulating strings. This includes a recursive definition of the replace-all function considered in this paper\footnote{\cite{TCJ16} used the notation \textbf{replace}}. The authors argue that ``the difficulty comes from ``recursively defined" functions such as replace, making state-of-the-art algorithms non-terminating." They proposed a progressive search algorithm, %to not only mitigate the problem of non-terminating reasoning but also guide the search towards a “minimal solution” when the input formula is in fact satisfiable. We have 
implemented within the state-of-the-art Z3 framework, with experimental evaluations. The algorithm is genetic and  applicable to all recursively defined string functions, but it is doomed to be incomplete as reasoning about unbounded strings defined recursively is in general an undecidable problem.   

The focus of our work is on the fundamental issue of decidability, and this is complementary to the work. Our result may be considered a completeness guarantee for existing string solver. 
}

%Importantly, we have enabled conflict clause learning for string theory so that our solver can be used effectively in the setting of program verification. Finally, our experimental evaluation shows leadership in a large benchmark suite, and a first deployment for another benchmark suite which requires reasoning about string formulas of a class that has not been solved before.



%===========================================================
We conclude with a discussion of related work and future work, focussing on
(1) decidability, and (2) heuristics and implementation.

\smallskip
\noindent
\textbf{Decidability.}
\emph{Length constraints} --- i.e. an assertion 
$\varphi(\Len(x_1),\ldots,\Len(x_n))$, where $\varphi$ is a Presburger formula
and $\Len(x_i)$ is an integer variable interpreted as the length of the string
$x_i$ --- have been studied in the context of string solving. It is
a major open problem whether the theory of concatenation with length
constraints is decidable \cite{Vijay-length}. Several extensions of this 
theory are undecidable (e.g. with letter counting \cite{buchi} and
string-number conversion \cite{GB16}).
Several decidable restrictions, however,
have been proposed including solved form \cite{Vijay-length} and acyclicity 
\cite{Abdulla14}, both of which (like straight-line constraint) impose 
syntactic restrictions on the way in which string equality can be
used in the constraints. It was shown in \cite{LB16} the decidability of path 
feasibility for symbolic
executions allowing \FT{} and concatenation in the 
assignments, and regular constraints, and length constraints 
in the assertions. If we allow the functions
$\replaceall_p(sub,rep)$ in the assignments (instead of
\FT{}/concatenation) and length constraints as assertions, path
feasibility becomes undecidable \cite{CCHLW18}. This also implies undecidability
of allowing length constraints in our constraint language with parametric
transducers. Fortunately, decidability can be easily recovered in some
cases. One such case is when the length constraints $\varphi$ has only
one string variable $x_1$, e.g., $\Len(x_1) > 7$. In this case,
$\varphi(\Len(x_1))$ can be turned into a regular constraint $x_1 \in L$ for
some $L$. [This is because the set of integer solutions is effectively
a finite union of arithmetic progressions $\bigcup_{i=1}^n (a_i + b_i\N)$
(where $a_i + b_i\N := \{ a_i + b_in : n \in \N\}$), and each 
$\Len(x_i) \in (a_i + b_i\N)$ is equivalent to the regular constraint 
$x_i \in \ialphabet^{a_i} (\ialphabet^{b_i})^*$.] 
\OMIT{
	The second case is when
	the length constraints only relate input variables. In this case, running
	our algorithm on the constraint (with length constraints ignored)
	results in a conjunction of regular constraints on the input variables. Each
	regular constraint $x \in L$ on an input variable $x$ yields a length
	constraint $L_x$ on $x$ of the form a finite union of arithmetic progressions 
	\cite{Kozen-automata}. We collect all the length constraints in a big
	conjunction and run a solver for integer linear arithmetic.
	%\cite{Vijay-length}, acyclicity \cite{Abdulla14,BFL13}, and straight-line
	%\cite{LB16}) 
	%on the shape of string constraints. 
}

The complexity of the theory of concatenation with regular constraints is
known to be \textsc{pspace}-complete \cite{Plandowski,J17}.
%, although the exact
%complexity for the theory of concatenation \emph{without} regular constraints is
%a long-standing open problem (only known to be in between NP and PSPACE). 
The complexity of the straight-line logic with concatenation, finite
transducers, and regular constraints is  \textsc{expspace}-complete \cite{LB16}. The same 
complexity holds when we swap finite transducers with replaceall 
\cite{CCHLW18}. For functions, our logic %with sweeping parameterised transducers
%as primitive operations 
strictly subsumes these two logics (e.g. since it can
also express string reverse) and has precisely the same complexity  \textsc{expspace}.
%The straight-line logic of \cite{LB16} extended with length constraints is still
%EXPSPACE-complete. 
One future research avenue is to identify \emph{larger subclasses} of
constraints with parametric transducers that are still solvable in the same 
complexity class.
%
%, for which we 
%only have a nonelementary upper bound (and EXPSPACE lower bound).

In this paper, we have combined two powerful formalisms 
(two-way finite transducers and replaceall) into a single formalism. Since we 
are considering
only two-way transducers that define functions, they are equivalent to 
two-way deterministic finite transducers, streaming transducers, and
MSO-definable transductions 
\cite{EH01,AC10,AD11}. On the other hand, one-way transducers that define 
functions are strictly more expressive than deterministic one-way transducers
\cite{Berstel}. 
\OMIT{
	In either case, it makes sense to consider nondeterministic
	transducers for a succinctness reason. %For future work, it 
}
\OMIT{
	It is also interesting to note that checking whether a \FFT{} defines a function
	is decidable \cite{CK87} (in fact, in polynomial-time for \FT{} \cite{GI83}),
	which implies decidability as well for \PPT{}.
}
%to consider 
%In this case, path feasibility of a
%symbolic execution $S$
%is defined in an \emph{angelic} way: each function in $S$
%should be able to produce \emph{some} output string that takes $S$ to the end of the
%program.
\OMIT{
	Note that the generic decision procedure in the proof of
	Theorem \ref{th:gen} still works when we admit relations, and the regularity
	conditions for \PPT{} and \PT{} are still satisfied whether or not they define
	functions. Therefore, our generic decision procedures would extend to ,
	albeit with a worse complexity. 
}

\OMIT{
	Symbolic automata were initially introduced in \cite{Watson96}, then
	investigated in \cite{NG01}, with motivation from natural language processing.
	The recent development of this topic was mainly driven by modern regular
	expression analysis and advanced web security analysis \cite{Vea13}, which
	require large alphabets (e.g. UTF16). Symbolic transducers were introduced in 
	\cite{symbolic-transducer}. See \cite{symbolic-transducer-power} for an 
	excellent survey. % on symbolic automata/transducers. 
}
%Prior to this work, symbolic 
%automata/transducers was never studied in the context of string solving.
%;
%the closest work related to this is the paper \cite{HJLRV18}, where the
%constraint language admits bitvectors in the alphabet.

%Symbolic automata/transducers were introduced by Veanes \emph{et al.}
%\cite{symbolic-transducer,symbolic-transducer-power}. \anthony{Say more}

%array theory? They can't concatenate and we can't do $a[i] = b[i]$.


\smallskip
\noindent
\textbf{Heuristics and implementation.}
Theoretical algorithms
(e.g. Makanin's algorithm \cite{Makanin}) typically do not directly lead to 
practical solvers. At the same time, the classes of constraints that are 
required in practice sometimes require string operations that are not covered by
decidable string constraint languages.
%(e.g.~length constraints, replaceall, 
%transducers, etc.). 
For these reasons,
there is a large amount of work on heuristics for developing practical
(often incomplete) string solvers, e.g., 
\cite{BTV09,Berkeley-JavaScript,HAMPI,Stranger,YABI14,Abdulla14,fang-yu-circuits,Abdulla17,HJLRV18,S3,TCJ16,Z3-str,Z3-str2,cvc4,Saner,RVG12}.
%Some incomplete heuristics were developed that could work in practice,
Some practical heuristics include
bounding string lengths (e.g. \cite{HAMPI,Berkeley-JavaScript,BTV09}), 
induction, overapproximations \cite{Stranger,YABI14}, interpolation
\cite{Abdulla14}, and flat automata \cite{Abdulla17}, to name a few. 
Focusing on semi-algorithms also allows highly expressive (but undecidable) 
string 
constraint languages, e.g., recursively defined functions \cite{S3,TCJ16}. 
Recently, the study of decidability of string constraints have also resulted in
automata-theoretic algorithms that are amenable to implementation, e.g.,
acyclic logic with concatenation, regular constraints, and length constraints 
\cite{Abdulla14}, and straight-line logic with finite transducers (or
replaceall), concatenation, and regular constraints 
\cite{HJLRV18,fang-yu-circuits,yan-tool}. 
%None of these techniques, however,
%would immediately lead to a fast string solver for string constraints with 
%parametric transducers as primitive operations.
%(e.g. global representations
%of solutions employed in \cite{HJLRV18} will not work for parametric
%transducers). 
We leave the development of practical heuristics for our more expressive 
constraint language for future work.

\OMIT{
	The development of theory and implementation of string solvers of course
	has to go hand in hand with the development of string constraint benchmarks.
}
We mention some interesting benchmarks that are available for 
string constraints. The first one is Kaluza 
benchmarks
\cite{Berkeley-JavaScript}, which contain
string constraints with concatenation, regular constraints, and length
constraints. It was shown in \cite{Vijay-length}
that almost all the constraints are already in
\emph{solved form} (in particular, also straight-line). Some of these length
constraints are also expressible as regular constraints.
The second is SLOG benchmarks \cite{fang-yu-circuits}, which contain
string constraints with concatenation, (a restricted class of) replaceall, and 
regular constraints. The third is SLOTH benchmarks \cite{HJLRV18}, which
contains string constraints with concatenation, finite transducers, and
regular constraints. Most of these benchmarking examples are expressible in our
decidable constraint language.
\OMIT{
	In this paper, we have described an application of
	string constraints for modelling string functions required for context-sensitive
	auto-sanitisation web templating \cite{SSS11}. An immediate future research 
	direction is to develop string constraint benchmarks based on this application.
}

%Some of these techniques utilise the power of
%alternating automata (over bitvectors)